### 2015-07-23 paper.tex
Added initial template

\documentclass{acm_proc_article-sp}

\begin{document}

\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms to
the formatting guidelines for ACM SIG Proceedings.
It complements the document \textit{Author's Guide to Preparing
ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}. This
source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning, clark:pct, braams:babel, herlihy:methodology},
conference
proceedings \cite{clark:pct} or books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

%\begin{figure}
%\centering
%\epsfig{file=fly.eps}
%\caption{A sample black and white graphic (.eps format).}
%\end{figure}
%
%\begin{figure}
%\centering
%\epsfig{file=fly.eps, height=1in, width=1in}
%\caption{A sample black and white graphic (.eps format)
%that has been resized with the \texttt{epsfig} command.}
%\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

%\begin{figure}
%\centering
%\psfig{file=rosette.ps, height=1in, width=1in,}
%\caption{A sample black and white graphic (.ps format) that has
%been resized with the \texttt{psfig} command.}
%\end{figure}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

%\begin{figure*}
%\centering
%\epsfig{file=flies.eps}
%\caption{A sample black and white graphic (.eps format)
%that needs to span two columns of text.}
%\end{figure*}
and don't forget to end the environment with
{figure*}, not {figure}!

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The acm\_proc\_article-sp document class file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
\balancecolumns
% That's all folks!
\end{document}

### 2015-08-10 paper.tex
Initial outline

@@
 \documentclass{acm_proc_article-sp}
+
+\usepackage{url}
 
@@
 
-\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
-Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
-\subtitle{[Extended Abstract]
-\titlenote{A full version of this paper is available as
-\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-\LaTeX$2_\epsilon$\ and BibTeX} at
-\texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
+\title{Experiences designing and building a PRAGMA Cloud Scheduler}
+\subtitle{[Short Paper]}
 
-\numberofauthors{8} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
+\numberofauthors{3}
+
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
+% 1st author      
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
+Shava Smallen\\
+      \affaddr{San Diego Supercomputer Center}\\
+      \affaddr{University of California San Diego}\\
+       \email{ssmallen@sdsc.edu}
 % 2nd. author
 \alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
-% 6th. author
-\alignauthor Charles Palmer\\
-       \affaddr{Palmer Research Laboratories}\\
-       \affaddr{8600 Datapoint Drive}\\
-       \affaddr{San Antonio, Texas 78229}\\
-       \email{cpalmer@prl.com}
+Nadya Williams\\
+      \affaddr{San Diego Supercomputer Center}\\
+      \affaddr{University of California San Diego}\\
+       \email{nadya@sdsc.edu}
+\and % 3rd. author
+\alignauthor Philip Papadopoulos \\
+      \affaddr{San Diego Supercomputer Center}\\
+      \affaddr{University of California San Diego}\\
+       \email{phil@sdsc.edu}
+%\and  % use '\and' if you need 'another row' of author names
 }
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{30 July 1999}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
 
@@
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms to
-the formatting guidelines for ACM SIG Proceedings.
-It complements the document \textit{Author's Guide to Preparing
-ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}. This
-source file has been written with the intention of being
-compiled under \LaTeX$2_\epsilon$\ and BibTeX.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and institutions from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.
+Over the past X years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
+During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for institutions to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.
+To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resource reservations.
+This short paper discusses the design process and initial prototype of a simple cloud scheduler for PRAGMA.
 
-The developers have tried to include every imaginable sort
-of ``bells and whistles", such as a subtitle, footnotes on
-title, subtitle and authors, as well as in the text, and
-every optional component (e.g. Acknowledgments, Additional
-Authors, Appendices), not to mention examples of
-equations, theorems, tables and figures.
-
-To make best use of this sample document, run it through \LaTeX\
-and BibTeX, and compare this source code with the printed
-output produced by the dvi file.
+%add contribute and leverage shared resources
 \end{abstract}
@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\terms{Theory}
+%terms{Cloud, Scheduling, Resource Sharing}
 
-\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings
+\keywords{Cloud, Scheduling, Resource Sharing} % NOT required for Proceedings
 
 \section{Introduction}
-The \textit{proceedings} are the records of a conference.
-ACM seeks to give these conference by-products a uniform,
-high-quality appearance.  To do this, ACM has some rigid
-requirements for the format of the proceedings documents: there
-is a specified format (balanced  double columns), a specified
-set of fonts (Arial or Helvetica and Times Roman) in
-certain specified sizes (for instance, 9 point for body copy),
-a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
-the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
-and (1.9 cm [.75"]) left and right; specified column width
-(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).
 
-The good news is, with only a handful of manual
-settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
-and {\texttt{\char'134 alignauthor}} commands, you have
-already used; another, {\texttt{\char'134 balancecolumns}}, will
-be used in your very last run of \LaTeX\ to ensure
-balanced column heights on the last page.}, the \LaTeX\ document
-class file handles all of this for you.
+some text~\cite{pragmaWeb}
 
-The remainder of this document is concerned with showing, in
-the context of an ``actual'' document, the \LaTeX\ commands
-specifically available for denoting the structure of a
-proceedings paper, rather than with giving rigorous descriptions
-or explanations of such commands.
+Repeat longer version of text in abstract with appropriate citations.
 
-\section{The {\secit Body} of The Paper}
-Typically, the body of a paper is organized
-into a hierarchical structure, with numbered or unnumbered
-headings for sections, subsections, sub-subsections, and even
-smaller sections.  The command \texttt{{\char'134}section} that
-precedes this paragraph is part of such a
-hierarchy.\footnote{This is the second footnote.  It
-starts a series of three footnotes that add nothing
-informational, but just give an idea of how footnotes work
-and look. It is a wordy one, just so you see
-how a longish one plays out.} \LaTeX\ handles the numbering
-and placement of these headings for you, when you use
-the appropriate heading commands around the titles
-of the headings.  If you want a sub-subsection or
-smaller part to be unnumbered in your output, simply append an
-asterisk to the command name.  Examples of both
-numbered and unnumbered headings will appear throughout the
-balance of this sample document.
+Discuss requirements and technologies like pragma boot, etc.
 
-Because the entire article is contained in
-the \textbf{document} environment, you can indicate the
-start of a new paragraph with a blank line in your
-input file; that is why this sentence forms a separate paragraph.
+% enable varying -- time limited -- not get bogged down by deployment of technologies
 
-\subsection{Type Changes and {\subsecit Special} Characters}
-We have already seen several typeface changes in this sample.  You
-can indicate italicized words or phrases in your text with
-the command \texttt{{\char'134}textit}; emboldening with the
-command \texttt{{\char'134}textbf}
-and typewriter-style (for instance, for computer code) with
-\texttt{{\char'134}texttt}.  But remember, you do not
-have to indicate typestyle changes when such changes are
-part of the \textit{structural} elements of your
-article; for instance, the heading of this subsection will
-be in a sans serif\footnote{A third footnote, here.
-Let's make this a rather short one to
-see how it looks.} typeface, but that is handled by the
-document class file. Take care with the use
-of\footnote{A fourth, and last, footnote.}
-the curly braces in typeface changes; they mark
-the beginning and end of
-the text that is to be in the different typeface.
+% view available resources and scientific application virtual machines
+\section{Scheduler Design}
 
-You can use whatever symbols, accented characters, or
-non-English characters you need anywhere in your document;
-you can find a complete list of what is
-available in the \textit{\LaTeX\
-User's Guide}\cite{Lamport:LaTeX}.
+Discuss process of narrowing down scheduler design from meeting ideas 
+(e.g., batch scheduler, Grid 5000, GENI, Google doc, DHCP leases).
 
-\subsection{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of
-the three are discussed in the next sections.
+\section{Calendaring systems}
 
-\subsubsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an
-inline or in-text formula.  It is produced by the
-\textbf{math} environment, which can be
-invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
-construction or with the short form \texttt{\$. . .\$}. You
-can use any of the symbols and structures,
-from $\alpha$ to $\omega$, available in
-\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
-few examples of in-text equations in context. Notice how
-this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
+Discuss selection of Booked.
 
-\subsubsection{Display Equations}
-A numbered display equation -- one set off by vertical space
-from the text and centered horizontally -- is produced
-by the \textbf{equation} environment. An unnumbered display
-equation is produced by the \textbf{displaymath} environment.
+\section{Cloud Scheduler Pilot}
 
-Again, in either environment, you can use any of the symbols
-and structures available in \LaTeX; this section will just
-give a couple of examples of display equations in context.
-First, consider the equation, shown as an inline equation above:
-\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
+Discuss creation of Cloud Scheduler Pilot.
 
-\subsection{Citations}
-Citations to articles \cite{bowman:reasoning, clark:pct, braams:babel, herlihy:methodology},
-conference
-proceedings \cite{clark:pct} or books \cite{salas:calculus, Lamport:LaTeX} listed
-in the Bibliography section of your
-article will occur throughout the text of your article.
-You should use BibTeX to automatically produce this bibliography;
-you simply need to insert one of several citation commands with
-a key of the item cited in the proper location in
-the \texttt{.tex} file \cite{Lamport:LaTeX}.
-The key is a short reference you invent to uniquely
-identify each work; in this sample document, the key is
-the first author's surname and a
-word from the title.  This identifying key is included
-with each item in the \texttt{.bib} file for your article.
+\section{Conclusions and Future Work}
 
-The details of the construction of the \texttt{.bib} file
-are beyond the scope of this sample document, but more
-information can be found in the \textit{Author's Guide},
-and exhaustive details in the \textit{\LaTeX\ User's
-Guide}\cite{Lamport:LaTeX}.
-
-This article shows only the plainest form
-of the citation command, using \texttt{{\char'134}cite}.
-This is what is stipulated in the SIGS style specifications.
-No other citation format is endorsed.
-
-\subsection{Tables}
-Because tables cannot be split across pages, the best
-placement for them is typically the top of the page
-nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and
-the table caption.  The contents of the table itself must go
-in the \textbf{tabular} environment, to
-be aligned properly in rows and columns, with the desired
-horizontal and vertical rules.  Again, detailed instructions
-on \textbf{tabular} material
-is found in the \textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table 1 is included in the input file; compare the
-placement of the table here with the table in the printed
-dvi output of this document.
-
-\begin{table}
-\centering
-\caption{Frequency of Special Characters}
-\begin{tabular}{|c|c|l|} \hline
-Non-English or Math&Frequency&Comments\\ \hline
-\O & 1 in 1,000& For Swedish names\\ \hline
-$\pi$ & 1 in 5& Common in math\\ \hline
-\$ & 4 in 5 & Used in business\\ \hline
-$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-\hline\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of
-the page's live area, use the environment
-\textbf{table*} to enclose the table's contents and
-the table caption.  As with a single-column table, this wide
-table will ``float" to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table 2 is included in the input file; again, it is
-instructive to compare the placement of the
-table here with the table in the printed dvi
-output of this document.
-
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-\subsection{Figures}
-Like tables, figures cannot be split across pages; the
-best placement for them
-is typically the top or the bottom of the page nearest
-their initial cite.  To ensure this proper ``floating'' placement
-of figures, use the environment
-\textbf{figure} to enclose the figure and its caption.
-
-This sample document contains examples of \textbf{.eps}
-and \textbf{.ps} files to be displayable with \LaTeX.  More
-details on each of these is found in the \textit{Author's Guide}.
-
-%\begin{figure}
-%\centering
-%\epsfig{file=fly.eps}
-%\caption{A sample black and white graphic (.eps format).}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\epsfig{file=fly.eps, height=1in, width=1in}
-%\caption{A sample black and white graphic (.eps format)
-%that has been resized with the \texttt{epsfig} command.}
-%\end{figure}
-
-As was the case with tables, you may want a figure
-that spans two columns.  To do this, and still to
-ensure proper ``floating'' placement of tables, use the environment
-\textbf{figure*} to enclose the figure and its caption.
-
-Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
-used; use
-the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
-commands as appropriate for the different file types.
-
-\subsection{Theorem-like Constructs}
-Other common constructs that may occur in your article are
-the forms for logical constructs like theorems, axioms,
-corollaries and proofs.  There are
-two forms, one produced by the
-command \texttt{{\char'134}newtheorem} and the
-other by the command \texttt{{\char'134}newdef}; perhaps
-the clearest and easiest way to distinguish them is
-to compare the two in the output of this sample document:
-
-This uses the \textbf{theorem} environment, created by
-the\linebreak\texttt{{\char'134}newtheorem} command:
-\newtheorem{theorem}{Theorem}
-\begin{theorem}
-Let $f$ be continuous on $[a,b]$.  If $G$ is
-an antiderivative for $f$ on $[a,b]$, then
-\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
-\end{theorem}
-
-The other uses the \textbf{definition} environment, created
-by the \texttt{{\char'134}newdef} command:
-\newdef{definition}{Definition}
-\begin{definition}
-If $z$ is irrational, then by $e^z$ we mean the
-unique number which has
-logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
-\end{definition}
-
-%\begin{figure}
-%\centering
-%\psfig{file=rosette.ps, height=1in, width=1in,}
-%\caption{A sample black and white graphic (.ps format) that has
-%been resized with the \texttt{psfig} command.}
-%\end{figure}
-
-Two lists of constructs that use one of these
-forms is given in the
-\textit{Author's  Guidelines}.
-
-%\begin{figure*}
-%\centering
-%\epsfig{file=flies.eps}
-%\caption{A sample black and white graphic (.eps format)
-%that needs to span two columns of text.}
-%\end{figure*}
-and don't forget to end the environment with
-{figure*}, not {figure}!
-
-There is one other similar construct environment, which is
-already set up
-for you; i.e. you must \textit{not} use
-a \texttt{{\char'134}newdef} command to
-create it: the \textbf{proof} environment.  Here
-is a example of its use:
-\begin{proof}
-Suppose on the contrary there exists a real number $L$ such that
-\begin{displaymath}
-\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-\end{displaymath}
-Then
-\begin{displaymath}
-l=\lim_{x\rightarrow c} f(x)
-= \lim_{x\rightarrow c}
-\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
-= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-\frac{f(x)}{g(x)} = 0\cdot L = 0,
-\end{displaymath}
-which contradicts our assumption that $l\neq 0$.
-\end{proof}
-
-Complete rules about using these environments and using the
-two different creation commands are in the
-\textit{Author's Guide}; please consult it for more
-detailed instructions.  If you need to use another construct,
-not listed therein, which you want to have the same
-formatting as the Theorem
-or the Definition\cite{salas:calculus} shown above,
-use the \texttt{{\char'134}newtheorem} or the
-\texttt{{\char'134}newdef} command,
-respectively, to create it.
-
-\subsection*{A {\secit Caveat} for the \TeX\ Expert}
-Because you have just been given permission to
-use the \texttt{{\char'134}newdef} command to create a
-new form, you might think you can
-use \TeX's \texttt{{\char'134}def} to create a
-new command: \textit{Please refrain from doing this!}
-Remember that your \LaTeX\ source code is primarily intended
-to create camera-ready copy, but may be converted
-to other forms -- e.g. HTML. If you inadvertently omit
-some or all of the \texttt{{\char'134}def}s recompilation will
-be, to say the least, problematic.
-
-\section{Conclusions}
-This paragraph will end the body of this sample document.
-Remember that you might still have Acknowledgments or
-Appendices; brief samples of these
-follow.  There is still the Bibliography to deal with; and
-we will make a disclaimer about that here: with the exception
-of the reference to the \LaTeX\ book, the citations in
-this paper are to articles which have nothing to
-do with the present subject and are used as
-examples only.
-%\end{document}  % This is where a 'short' article might terminate
+Some interesting conclusion.
 
 %ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
+%\section{Acknowledgments}
 
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
 \bibliographystyle{abbrv}
-\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-%\balancecolumns
-\appendix
-%Appendix A
-\section{Headings in Appendices}
-The rules about hierarchical headings discussed above for
-the body of the article are different in the appendices.
-In the \textbf{appendix} environment, the command
-\textbf{section} is used to
-indicate the start of each Appendix, with alphabetic order
-designation (i.e. the first is A, the second B, etc.) and
-a title (if you include one).  So, if you need
-hierarchical structure
-\textit{within} an Appendix, start with \textbf{subsection} as the
-highest level. Here is an outline of the body of this
-document in Appendix-appropriate form:
-\subsection{Introduction}
-\subsection{The Body of the Paper}
-\subsubsection{Type Changes and  Special Characters}
-\subsubsection{Math Equations}
-\paragraph{Inline (In-text) Equations}
-\paragraph{Display Equations}
-\subsubsection{Citations}
-\subsubsection{Tables}
-\subsubsection{Figures}
-\subsubsection{Theorem-like Constructs}
-\subsubsection*{A Caveat for the \TeX\ Expert}
-\subsection{Conclusions}
-\subsection{Acknowledgments}
-\subsection{Additional Authors}
-This section is inserted by \LaTeX; you do not insert it.
-You just add the names and information in the
-\texttt{{\char'134}additionalauthors} command at the start
-of the document.
-\subsection{References}
-Generated by bibtex from your ~.bib file.  Run latex,
-then bibtex, then latex twice (to resolve references)
-to create the ~.bbl file.  Insert that ~.bbl file into
-the .tex source file and comment out
-the command \texttt{{\char'134}thebibliography}.
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-\section{More Help for the Hardy}
-The acm\_proc\_article-sp document class file itself is chock-full of succinct
-and helpful comments.  If you consider yourself a moderately
-experienced to expert user of \LaTeX, you may find reading
-it useful but please remember not to change it.
-\balancecolumns
-% That's all folks!
+\bibliography{paper}
+
 \end{document}

### 2015-08-12 paper.tex
Added intro

@@
 \begin{abstract}
-The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and institutions from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.
 Over the past X years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
-During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for institutions to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.
+During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.
 To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resource reservations.
@@
 
-some text~\cite{pragmaWeb}
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
+Founded in 2002, PRAGMA originally sought to advance the use of grid technologies in applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.
+This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and conduct scientific experiments.
+Testbed sites needed to install at a minimum Globus~\cite{globus} and a local HPC job scheduler as well as other optional software such as MPICH-G2~\cite{mpichg2} and Ninf-G~\cite{ninfg}.
+By November 2006, nineteen site in thirteen countries were part of the testbed, with a total of 662 CPUs, nearly 1 terabyte of memory, and 7.3 terabytes of online storage~\cite{pragmaReport2006}.
+In 2009, the testbed had grown to twenty-seven sites in fifteen countries with a total of 1008 CPUs, more than 1.3 terabytes of memory, and over 24.7 terabytes of online storage.
+However by 2011, the number of sites started to decline and maintaining the numerous and complex middleware and scientific applications at a local site required significant people effort and expertise.
+Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and lower the amount of effort an site would need to participate in the testbed.
+
+The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
+One early example was an AutoDoc~\cite{autodock} virtual machine created for the Avian Flu research team in 2011~\cite{pragmaReport2011}.
+By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.
+To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus to the creation and management of virtual clusters in 2013.
+The following technologies have been explored to facilitate both the operation and use of virtual clusters on the PRAGMA cloud testbed.
+
+\textbf{pragma\_boot:}  Rather than requiring a single cloud system, PRAGMA sites are free to  deploy any cloud tool that best fits their expertise.
+However, enabling a virtual cluster to be ported to different cloud systems is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.
+In 2013, three pilot sites (UCSD, AIST, NCHC) developed an automated virtual cluster porting script and demonstrated the same virtual cluster image being booted in three different Cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
+Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} and "drivers" were implemented for both Rocks and OpenNebula.
+
+\textbf{Personal Cloud Controller:}  To provide users with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC) tool was started in 2014.
+This lightweight tool was designed to manage startup,  status monitoring, and shutdown  of a virtual cluster and was built on top of pragma\_boot and a well-known resource management tool called HTCondor~\cite{condor}.
+PCC also provided an option to create a multi-site virtual cluster leveraging a network overlay tool called IPOP~\cite{ipop} for the private network.
+
+\textbf{Software-defined Networking}:   PRAGMA began investigating software-defined networking technologies like OpenFlow~\cite{openflow} in 2013 as a way to create a private network between multi-site virtual clusters and to protect access to sensitive datasets~\cite{pragmaReport2013}.
+PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
+In 2014, The PRAGMA-ENT team worked to create an international reliable direct point-to-point data Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} to provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
+
+In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate access to the above technologies.
+The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.
+A summary of calendar systems that were considered is provided in Section~\ref{Sec:Calendars}.
+Section~\ref{Sec:Pilot} discusses our early implementation with one of the calendar systems called Booked and Section~\ref{Sec:Conclusions} concludes with our planned future work.
+
+% NOTES AND SNIPPET TEXT THAT WAS NO LONGER RELEVANT
+%In 2012, a \textit{vm-deploy} script was developed to automate the process of launching a virtual machine from distributed Gfarm filesystem~\cite{gfarm) to an OpenNebula deployment.
+, which later became \textit{pragma\_boot} in 2013, to automate the VM porting process from one cloud deployment to another~\cite{pragmaReport2012,pragmaReport2013}.
+
+%In Phase 2, PRAGMA started using Gfarm~\cite{gfarm} as a mechanism to distribute and migrate VMs to different sites~\cite{pragmaReport2012} and in Phase 3 grew the cloud testbed to ten sites in 2012 with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.
+To facilitate the creation of application VMs, a web-based, easy-to-use user interface was developed for the Ezilla tool~\cite{ezilla} so that users can drag-n-drop the applications they need into their own custom VM image~\cite{pragmaReport2013}.
+
+% 2005-2006: 19 sites in 13 countries, with a total of 662 CPUs, near- ly 1 terabyte of memory, and 7.3 terabytes of online storage.
+% 2006-2007: The PRAGMA testbed grew from 19 sites in 13 countries to 26 sites in 14 countries, with a total of 726 CPUs, more than half a terabyte of memory, and 13.2 terabytes of online storage
+% 2007-2008: During this last year, the PRAGMA grid grew from 26 sites in 14 countries/regions to 30 sites in 17 countries/regions, with a total of 1109 CPUs, more than one terabyte of memory, and 26.7 terabytes of online storage.
+% 2008-2009: During the last year, the PRAGMA Grid has grown to 27 sites in 15 countries/regions with total of 1008 CPUs, more than 1.3 ter- abytes of memory, and over 24.7 terabytes of online storage.
+% 2009-2010:  Currently, the PRAGMA Grid has 24 sites in 16 countries/regions which provides a total of 1022 CPUs, more than 1.3 terabytes of memory, and over 24.8 terabytes of online storage. 
+% 2010-2011: Since PRAGMA 17 (October 2009, Hanoi), the Resources Working Group has decided to migrate from grid to cloud through experimentation with virtualization technologies
+% 2011-2012: At PRAGMA 20 (March 2011, Hong Kong), the three pilot sites demonstrated this Phase 1 experiment and findings.
+These results have excited many PRAGMA sites and motivated them to join this effort.
+Since then, seven more sites have set up VM hosting services and migrated the three application VMs.
+These sites are: Indiana University (IU), ASTI, MIMOS, LZU, Osaka University, CNIC and University of Hyderabad (UoHyd).
+% 2012-2013: Our progress in porting (a form of sharing) VM images has taken place in three phases.
+In the first phase, (completed in early 2011), we successfully demonstrated a manual port of three application VM images among three different VM hosting environments (i.e., a pairing of a VM hosting platform with VM hosting managing software, e.g.
+Open Nebula).
+Phase 2 began at PRAGMA 20 (March 2011, Hong Kong), where, based on what we learned from the phase 1 experiments, we designed a PRAGMA cloud infrastructure to use Gfarm for VM image depository and sharing and to automate VM deployment process on various virtualization platforms that make up the PRAGMA multi-cloud.
+At PRAGMA 21 (October 2011, Japan), we demonstrated the automated deployment of a GEO Grid VM image from Gfarm to three pilot sites?AIST, NCHC and UCSD.
+After PRAGMA 21, we started Phase 3 of the VM sharing experiment with four objectives: 1) to expand PRAGMA cloud resources; 2) to enhance Gfarm functionality and performance; 3) to author and run more application VMs and 4) to develop an easy-to-use user interface for creating VM images.
+% 2013-2014:
+
+\section{Scheduler Requirements}
+
+Sites willing to participate will have to install a PRAGMA Package - a very small set of components.
+Users can request access to resources similar to requesting a room (i.e., a room reservation system)
+PCC would be used to automatically start and monitor user's VMs slash VC via pragma\_boot up at sites 
+Only support 10s of users
 
@@
 \section{Scheduler Design}
+\label{Sec:Design}
 
@@
 \section{Calendaring systems}
+\label{Sec:Calendars}
 
@@
 \section{Cloud Scheduler Pilot}
+\label{Sec:Pilot}
 
@@
 \section{Conclusions and Future Work}
+\label{Sec:Conclusions}
 

### 2015-08-13 paper.tex
Revised intro slightly and added requirements

@@
 The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.
-Over the past X years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
+Over the past four years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
 During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.
@@
 By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.
-To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus to the creation and management of virtual clusters in 2013.
+In addition to the previously managed cloud tools, To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus to the creation and management of virtual clusters in 2013.
 The following technologies have been explored to facilitate both the operation and use of virtual clusters on the PRAGMA cloud testbed.
@@
 
-In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate access to the above technologies.
+In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate running on the different cloud deployments and leverage the above technologies.
 The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.
@@
 
-Sites willing to participate will have to install a PRAGMA Package - a very small set of components.
+PRAGMA had the following three key requirements for a cloud testbed scheduler:
+
+\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it be lightweight and require  minimal effort and minimal expertise for a site to add their cloud deployment to the list of schedulable resources.
+Similarly the scheduler would need to work across multiple cloud deployment tools (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install one specific cloud deployment tool.
+Sites  just need to run a SSH service and install the pragma\_boot package, which is a small Python package and can be installed via RPM and can be configured in less than an hour.
+
+\textbf{Easy to use}:  To execute a virtual cluster on the PRAGMA cloud testbed currently, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and if pragma\_boot is not available, perform all the steps manually to launch their virtual cluster.
+The design of the PRAGMA cloud scheduler should first simplify that process by ensuring pragma\_boot gets deployed more broadly and ported it to more cloud systems as needed.
+The scheduler should also provide a simple Web interface for users to see the available resources at sites, construct their virtual cluster, and manage their images.
+
+\textbf{Scale to tens of users:}  While the PRAGMA community was composed of nineteen active institution sites in 2014, the number of expected users for the cloud testbed is in the tens of users, not hundreds nor thousands.
+Therefore, our scheduler can be simplified and we can give high priority to the requirements of low participation overhead and easy of use.
+
+% enable varying -- time limited -- not get bogged down by deployment of technologies
+
+% view available resources and scientific application virtual machines
+\section{Scheduler Design}
+\label{Sec:Design}
+
 Users can request access to resources similar to requesting a room (i.e., a room reservation system)
@@
 
-% enable varying -- time limited -- not get bogged down by deployment of technologies
-
-% view available resources and scientific application virtual machines
-\section{Scheduler Design}
-\label{Sec:Design}
+ENT
+one controller at participating site
+way to slice switch? specialized controller? 
+AutoVFlow by Hiroaki Yamanaka (NICT)
+For users interested in running VMs, cloud middleware (OpenStack, CloudStack, etc) should take care of the slicing.
+ Recurring problem: users tend to forget and leave idle VMs running.
+For users interested in cloud middleware research/development, and/or access to bare-metal hardware (e.g., ENT), scheduling of resources and coordination among users will be needed.
+Network slicing technologies (Auto-VFlow, FlowVisor, FlowSpace Firewall, etc) when deployed could help
+Use PBS as a scheduler (or Condor or Slurm?)
+Nimbus have a module that interfaces with PBS
+ Use PBS to schedule interactive sessions on bare metal
+Use a shared calendar so that users can make resource reservations (`a la Grid?5000)
+ORCA from GENI/Planetlab? 
+Leases
 
@@
 %ACKNOWLEDGMENTS are optional
-%\section{Acknowledgments}
+\section{Acknowledgments}
+
+Thank Mauricio and Jose.
 

### 2015-08-13 paper.tex
Latest version.

@@
 
-PRAGMA had the following three key requirements for a cloud testbed scheduler:
+PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users access to the resources at a time:
 
@@
 \textbf{Scale to tens of users:}  While the PRAGMA community was composed of nineteen active institution sites in 2014, the number of expected users for the cloud testbed is in the tens of users, not hundreds nor thousands.
-Therefore, our scheduler can be simplified and we can give high priority to the requirements of low participation overhead and easy of use.
+Therefore, we can prioritize simplicity over scalability and give higher priority to the requirements of low participation overhead and easy of use.
 
-% enable varying -- time limited -- not get bogged down by deployment of technologies
-
-% view available resources and scientific application virtual machines
 \section{Scheduler Design}
@@
 
-Users can request access to resources similar to requesting a room (i.e., a room reservation system)
-PCC would be used to automatically start and monitor user's VMs slash VC via pragma\_boot up at sites 
-Only support 10s of users
+In our design discussions, we first looked at the below existing solutions that could be adapted for PRAGMA's scheduling needs:
 
-Repeat longer version of text in abstract with appropriate citations.
+\textbf{Existing HPC schedulers}:  Open source batch schedulers like Slurm~\cite{slurm}, TORQUE~\cite{torque}, and HTCondor are commonly used in  HPC to manage space shared resources.
+While these batch schedulers  are widely used, they either do not work across multiple sites (TORQUE) or are too heavy weight for what PRAGMA needs, requiring custom configuration and multiple daemons at each site (Slurm and HTCondor).
+Furthermore, PRAGMA users change over time and we wanted a low barrier to entry like a simple Web based interface for scheduling, rather than the command-line tools and queue abstractions that these batch schedulers provide.
 
-Discuss requirements and technologies like pragma boot, etc.
+\textbf{Grid'5000:}   NEEDS TO BE WRITTEN.
+Add short description of Grid'5000.
+They use OAR which is a batch scheduler similar to above and the calendar gui called GridPrems is one of three options for interacting with Grid'5000.
+A key difference is that testbed resources are dedicated rather than like on PRAGMA where they could be shared;  I believe also single admin model.
+Interesting set of tools.
+We can say we liked their calendar approach the best.
 
-ENT
-one controller at participating site
-way to slice switch? specialized controller? 
-AutoVFlow by Hiroaki Yamanaka (NICT)
-For users interested in running VMs, cloud middleware (OpenStack, CloudStack, etc) should take care of the slicing.
- Recurring problem: users tend to forget and leave idle VMs running.
-For users interested in cloud middleware research/development, and/or access to bare-metal hardware (e.g., ENT), scheduling of resources and coordination among users will be needed.
-Network slicing technologies (Auto-VFlow, FlowVisor, FlowSpace Firewall, etc) when deployed could help
-Use PBS as a scheduler (or Condor or Slurm?)
-Nimbus have a module that interfaces with PBS
- Use PBS to schedule interactive sessions on bare metal
-Use a shared calendar so that users can make resource reservations (`a la Grid?5000)
-ORCA from GENI/Planetlab? 
-Leases
+\textbf{DHCP Leases:}  NEEDS TO BE WRITTEN.
 
-Discuss process of narrowing down scheduler design from meeting ideas 
-(e.g., batch scheduler, Grid 5000, GENI, Google doc, DHCP leases).
+\textbf{GENI/PlanetLab}:  NEEDS TO BE WRITTEN.  ORCA
 
-\section{Calendaring systems}
+Out of these approaches, we liked the Grid'5000 calendar reservation interface, GridPrems, for scheduling the best.
+The calendar reservation scheduling is an interface that all users are familiar with in their everyday lives and would be well suited to the different PRAGMA users who have varying levels of cloud expertise.
+
+\section{Calendar systems}
 \label{Sec:Calendars}
+
+In addition to GridPrems interface, we decided to look at other calendar reservation systems.
 
@@
 
-Thank Mauricio and Jose.
+The authors would like to thank Jose Fortes and Mauricio Tsugawa for their help during our initial design discussions and for providing resources at University of Florida during development.
 

### 2015-08-13 paper.tex
small typos and minor rewrites.

@@
 \begin{abstract}
-The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like biodiversity and lake ecology.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
 Over the past four years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
-During PRAGMA 27, the Resources Working group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and overlay networks.
-To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resource reservations.
+During PRAGMA 27 meeting, the Resources Working Group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and virtual network overlays (i.e.
+IPOP and ViNe).
+To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resources reservations.
 This short paper discusses the design process and initial prototype of a simple cloud scheduler for PRAGMA.
@@
 
-\keywords{Cloud, Scheduling, Resource Sharing} % NOT required for Proceedings
+\keywords{Cloud, Scheduling, Resource Sharing, Virtualization} % NOT required for Proceedings
 
@@
 
-The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of individuals and sites from around the Pacific Rim that actively collaborate to enable scientific expeditions in areas like computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
+PRAGMA researches actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
 Founded in 2002, PRAGMA originally sought to advance the use of grid technologies in applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.
-This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and conduct scientific experiments.
+This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and to conduct scientific experiments.
 Testbed sites needed to install at a minimum Globus~\cite{globus} and a local HPC job scheduler as well as other optional software such as MPICH-G2~\cite{mpichg2} and Ninf-G~\cite{ninfg}.
-By November 2006, nineteen site in thirteen countries were part of the testbed, with a total of 662 CPUs, nearly 1 terabyte of memory, and 7.3 terabytes of online storage~\cite{pragmaReport2006}.
-In 2009, the testbed had grown to twenty-seven sites in fifteen countries with a total of 1008 CPUs, more than 1.3 terabytes of memory, and over 24.7 terabytes of online storage.
-However by 2011, the number of sites started to decline and maintaining the numerous and complex middleware and scientific applications at a local site required significant people effort and expertise.
-Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and lower the amount of effort an site would need to participate in the testbed.
+In November 2006, nineteen sites in thirteen countries were part of the testbed.
+By 2009, the testbed had grown to twenty-seven sites in fifteen countries with a total of 1008 CPUs, more than 1.3 terabytes of memory, and over 24.7 terabytes of online storage.
+However, as of 2011, the number of sites started to decline and maintaining the numerous and complex middleware and scientific applications at each local site required significant people effort and expertise.
+Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and to lower the amount of effort any site would need to participate in the testbed.
 
-The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
-One early example was an AutoDoc~\cite{autodock} virtual machine created for the Avian Flu research team in 2011~\cite{pragmaReport2011}.
+The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
+One early example was an AutoDock~\cite{autodock} virtual cluster created for the Avian Flu Grid research team in 2011~\cite{pragmaReport2011}.
 By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.
-In addition to the previously managed cloud tools, To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus to the creation and management of virtual clusters in 2013.
-The following technologies have been explored to facilitate both the operation and use of virtual clusters on the PRAGMA cloud testbed.
+In 2013 PRAGMA shifted its  focus to the creation and management of virtual clusters to make it easier for users to assemble a multi-node virtual environment for running their scientific experiments.
+The following technologies have been explored to facilitate the operation and use of virtual clusters on the PRAGMA cloud testbed.
 
-\textbf{pragma\_boot:}  Rather than requiring a single cloud system, PRAGMA sites are free to  deploy any cloud tool that best fits their expertise.
-However, enabling a virtual cluster to be ported to different cloud systems is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.
-In 2013, three pilot sites (UCSD, AIST, NCHC) developed an automated virtual cluster porting script and demonstrated the same virtual cluster image being booted in three different Cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
-Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} and "drivers" were implemented for both Rocks and OpenNebula.
+\textbf{Pragma\_boot:}  rather than requiring a single cloud system at all sites, PRAGMA encourages the sites to  deploy any virtualization technology that best fits their expertise.
+However, enabling a virtual cluster to be ported to different cloud hosting environment is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.
+In 2013, three pilot sites (UCSD, AIST, NCHC) developed a script for an automated virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
+Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.
 
-\textbf{Personal Cloud Controller:}  To provide users with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC) tool was started in 2014.
-This lightweight tool was designed to manage startup,  status monitoring, and shutdown  of a virtual cluster and was built on top of pragma\_boot and a well-known resource management tool called HTCondor~\cite{condor}.
-PCC also provided an option to create a multi-site virtual cluster leveraging a network overlay tool called IPOP~\cite{ipop} for the private network.
+\textbf{Personal Cloud Controller:}  to provide users with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014.
+This lightweight tool was designed to manage startup, status monitoring, and shutdown of a virtual cluster and was built on top of pragma\_boot and a well-known resource and workload management system HTCondor~\cite{condor}.
+PCC also provided an option to create a multi-site virtual cluster leveraging an open-source virtual network overlay software IPOP~\cite{ipop} for creating its private network.
 
-\textbf{Software-defined Networking}:   PRAGMA began investigating software-defined networking technologies like OpenFlow~\cite{openflow} in 2013 as a way to create a private network between multi-site virtual clusters and to protect access to sensitive datasets~\cite{pragmaReport2013}.
+\textbf{Software-Defined Networking}:   PRAGMA began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network between multi-site virtual clusters and to protect access to sensitive datasets~\cite{pragmaReport2013}.
 PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
-In 2014, The PRAGMA-ENT team worked to create an international reliable direct point-to-point data Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} to provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
+In 2014, The PRAGMA-ENT team worked to create an international reliable direct point-to-point data Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that allows to provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
 
-In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate running on the different cloud deployments and leverage the above technologies.
+In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate virtual hosts and clusters running on the different cloud deployments via leverage the above described technologies.
 The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.
-A summary of calendar systems that were considered is provided in Section~\ref{Sec:Calendars}.
+A summary of calendar systems that were examined is provided in Section~\ref{Sec:Calendars}.
 Section~\ref{Sec:Pilot} discusses our early implementation with one of the calendar systems called Booked and Section~\ref{Sec:Conclusions} concludes with our planned future work.
@@
 
-PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users access to the resources at a time:
+PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users access to the resources at a  given time:
 
-\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it be lightweight and require  minimal effort and minimal expertise for a site to add their cloud deployment to the list of schedulable resources.
-Similarly the scheduler would need to work across multiple cloud deployment tools (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install one specific cloud deployment tool.
-Sites  just need to run a SSH service and install the pragma\_boot package, which is a small Python package and can be installed via RPM and can be configured in less than an hour.
+\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it had to be lightweight and required  minimal effort and minimal expertise for a site to add their cloud deployment to the list of resources available for scheduling.
+Similarly, the scheduler would need to work across multiple cloud deployments (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install specific cloud deployment tools in addition to their currently used.
 
-\textbf{Easy to use}:  To execute a virtual cluster on the PRAGMA cloud testbed currently, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and if pragma\_boot is not available, perform all the steps manually to launch their virtual cluster.
-The design of the PRAGMA cloud scheduler should first simplify that process by ensuring pragma\_boot gets deployed more broadly and ported it to more cloud systems as needed.
-The scheduler should also provide a simple Web interface for users to see the available resources at sites, construct their virtual cluster, and manage their images.
+NOTE THE FOLLOWING IS A PART OF AN IMPLEMENTATION NOT A REQUIREMENT, MOVE TO DIFFERENT SECTION THAT DESCRIBES IMPLEMENTATION: Sites  just need to run a SSH service and install the pragma\_boot package, which is a small Python package and can be installed via RPM and can be configured in less than an hour.
 
-\textbf{Scale to tens of users:}  While the PRAGMA community was composed of nineteen active institution sites in 2014, the number of expected users for the cloud testbed is in the tens of users, not hundreds nor thousands.
-Therefore, we can prioritize simplicity over scalability and give higher priority to the requirements of low participation overhead and easy of use.
+\textbf{Easy to use}:  Currently, to execute a virtual cluster on the PRAGMA cloud testbed, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and launch their virtual cluster either using pragma\_boot  if it is available or manually via a sequence of defined steps.
+The design of the PRAGMA cloud scheduler should first simplify this process by ensuring that pragma\_boot gets deployed more broadly and port it to more cloud environments as needed.
+The scheduler should also provide a simple Web interface for users to see the available resources and sites, construct their virtual cluster, and manage their images.
+
+\textbf{Scale to tens of users:}  While the PRAGMA community was composed of nineteen active institution sites in 2014, the initial number of expected users for the cloud testbed is in the tens of users, not hundreds nor thousands.
+Therefore, we can prioritize simplicity over scalability and give higher priority to the requirements of low participation overhead and ease of use.
 
@@
 
-In our design discussions, we first looked at the below existing solutions that could be adapted for PRAGMA's scheduling needs:
+In our design discussions, we first looked at the existing solutions that could be adapted for PRAGMA's scheduling needs:
 
-\textbf{Existing HPC schedulers}:  Open source batch schedulers like Slurm~\cite{slurm}, TORQUE~\cite{torque}, and HTCondor are commonly used in  HPC to manage space shared resources.
+\textbf{Existing HPC schedulers}:  Open source batch schedulers like Slurm~\cite{slurm}, TORQUE~\cite{torque}, and HTCondor are commonly used in  HPC to manage distributed  shared resources.
 While these batch schedulers  are widely used, they either do not work across multiple sites (TORQUE) or are too heavy weight for what PRAGMA needs, requiring custom configuration and multiple daemons at each site (Slurm and HTCondor).
-Furthermore, PRAGMA users change over time and we wanted a low barrier to entry like a simple Web based interface for scheduling, rather than the command-line tools and queue abstractions that these batch schedulers provide.
+Furthermore, to be adopted by PRAGMA users who are scientists in many different domains,  we wanted a low barrier to entry like a simple Web based interface for scheduling, rather than the command-line tools and queue abstractions that these batch schedulers provide.
 

### 2015-08-14 paper.tex
Started on calendar section.  Before merging Nadya's changes

--- 2015-08-13 paper.tex
@@
 
-In addition to GridPrems interface, we decided to look at other calendar reservation systems.
+Since we were primarily interested in a scheduler with a Web based calendar reservation interface like GridPrems, we decided to look at other open source calendar reservation systems that were available.
+We looked at several systems used primarily for tracking room and/or equipment reservations and evaluated the following criteria:
+
+\textbf{Open Source and customizable:}  Open source software is generally preferable because it provides it is free to use and can be customized as needed.
+Since we did not know of any existing Web interfaces for managing virtual clusters, we limited our search to customizable Open Source tools and evaluated how easy it would be to manage resources and users as well as add new fields and features.
+
+\textbf{Nice GUI interface:}  We evaluated the GUI interfaces of several calendar reservation tools and looked for those that had intuitive menus and navigation and  a clean and uncluttered look.
+Some of the GUI interfaces we looked at just used basic HTML tags and looked very clunky so we limited our search to tools that leveraged CSS and Javascript and looked more professional.
+
+\textbf{Installation and maintenance:}  We downloaded a number of tools and reviewed their installation instructions to see how easy they were to setup a demo instance.
+Most of the tools we looked at were developed in PHP and had a relational database backend that was could be served from a basic  Apache web server\~cite{apache} setup.
+A few of the tools were plugins for Drupal.
+However, we eliminated a number of tools because they had either too little documentation or complex installation instructions with several manual steps.
+
+We searched for candidate reservation tools using Google search and also searching specifically on Github and SourceForge.  The above criteria filtered out several tools.
+
+One category of tools we looked at were plugins to existing frameworks like Drupal~\cite{drupal} and Wordpress\~cite{wordpress}.
+We looked at Booking_calendar for Wordpress as well as Merci and Reservations for Drupal
+One category of tools that we looked at were developed in PHP and had a relational backend database such as Postgres an MySQL.
+We looked at Meeting Room Booking System on SourceForge~\cite{mrbs}, Booked from Twinkle Toes Software, and Classroombookings on Github,
 
 Discuss selection of Booked.
+
+Open source
+Easy to setup
+Nice GUI interface
+Report features
+REST API
+Customizable-ish
+LDAP and Active Directory support.
+Fine tuned roles and permissions.
+User and group quotas.
+
+Can only handle one reservation per resource at a time
+PHP changes can be painful (heavy OO makes it hard to find right files)
+Doc is sparse
+
+http://mrbs.sourceforge.net
+	PHP, Postgres/MySQL
+	Demo was a bit slow and a bit clunky looking
+Booked - http://www.twinkletoessoftware.com/products
+	Nice interface, lot of customization features
+	PHP, MySQL
+	RESTFUL API
+http://classroombookings.com/
+	Nice interface
+	PHP, MySQL
+http://www.unitime.org/unitime\_intro.php
+	Geeky looking, might have some good algorithms
+	Interface not as smooth
+	http://journal.code4lib.org/articles/2941
+	Clunky looking
+http://www.drupalrooms.com
+	More for hotel web sites ? e.g., prices, etc.
+https://groups.drupal.org/node/137544
+	Merci - https://www.drupal.org/project/merci
+		For equipment, etc.
+	https://www.drupal.org/project/reservations
+		Also seems pretty flexible
+
+ tried a few tools  here is the summary:
+1. OTA hotel management   http://rocks-86.sdsc.edu/hotelmis/setup/initialsetup.php
+    appears to be just for the hotel, no easy login, no easy change. Not useful
+2. Mybooking - in Malay and after changing the language and trying enenglishigsh i did not get far. -- hotel
+3. Booking_calendar - strictly room booking, very small in features, too much to rewrite. -- wordpress plugin
+4 Optaplanner appears to be a java app that runs outside of web. I tried examples and i am
+   not sure of its usability. The examples are running producing sometimes some output sometimes noe
+   and they are just java apps on some data. The examples  are not impressive even in terms of giving
+   an idea how things are run.
+
+A few more links are on http://opensource.com/business/15/1/top-project-management-tools-2015
+Started looking at  one of them, Tuleap, may be too big for what we need.
 

### 2015-08-14 paper.tex
Merged copy

@@
 \begin{abstract}
-The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
+PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
 Over the past four years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
@@
 
-PRAGMA researches actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) researches actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
 Founded in 2002, PRAGMA originally sought to advance the use of grid technologies in applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.
@@
 One category of tools we looked at were plugins to existing frameworks like Drupal~\cite{drupal} and Wordpress\~cite{wordpress}.
-We looked at Booking_calendar for Wordpress as well as Merci and Reservations for Drupal
+We looked at Booking\_calendar for Wordpress as well as Merci and Reservations for Drupal
 One category of tools that we looked at were developed in PHP and had a relational backend database such as Postgres an MySQL.
@@
 2. Mybooking - in Malay and after changing the language and trying enenglishigsh i did not get far. -- hotel
-3. Booking_calendar - strictly room booking, very small in features, too much to rewrite. -- wordpress plugin
+3. Booking\_calendar - strictly room booking, very small in features, too much to rewrite. -- wordpress plugin
 4 Optaplanner appears to be a java app that runs outside of web. I tried examples and i am

### 2015-08-14 paper.tex
Filled in pilot section with some text

@@
 \section{Introduction}
+\label{Sec:intro}
 
@@
 Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.
+In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront~\cite{cloudfront}.
 
@@
 Similarly, the scheduler would need to work across multiple cloud deployments (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install specific cloud deployment tools in addition to their currently used.
-
-NOTE THE FOLLOWING IS A PART OF AN IMPLEMENTATION NOT A REQUIREMENT, MOVE TO DIFFERENT SECTION THAT DESCRIBES IMPLEMENTATION: Sites  just need to run a SSH service and install the pragma\_boot package, which is a small Python package and can be installed via RPM and can be configured in less than an hour.
 
@@
 
+One key requirement for our cloud scheduler was to enable users to start and manage virtual clusters on the PRAGMA cloud testbed with minimal effort. 
+The basic workflow of how a user would start a virtual cluster and the steps are summarized below:
+
+\begin{enumerate}
+\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (i.e., IPOP, ViNE, or PRAGMA-ENT) that is needed.  
+\item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
+\item The user selects the subset of resources they want, the timeframe, virtual cluster image, and configuration details and submits it to the Cloud scheduler.
+Once the reservation is verified and confirmed, the user will get an email confirmation.
+\item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when it's ready for them to login.
+\end{enumerate}
+
+Figure~\ref{Fig:Flow} shows the above steps in a use case sequence diagram where \textit{PRAGMA Booking} refers to our customized Booked instance.
+PCC will monitor the health of the virtual cluster and when the reservation is close to expiring, will notify the user in case they want to extend their reservation.
+Otherwise, when the time expires or the user chooses to end their reservation, the virtual cluster will be shut down.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/flow.pdf}
+\caption{The flow of a user request to start a virtual cluster and the components that are involved in the cloud scheduler.}
+\label{Fig:Flow}
+\end{center}
+\end{figure}
+
+The other key requirement for the cloud scheduler was that it should require minimal effort for a PRAGMA sites to participate in the PRAGMA testbed.
+Therefore, we opted for a simple client-server architecture, where the client is a small PRAGMA package that interfaces to a site's local cloud system.
+Currently, the PRAGMA Package is just a SSH service and the pragma\_boot package, which is a small Python package that can be installed via RPM and can be configured in less than an hour.
+The server component is composed of the PRAGMA Booking interface and the PCC which gets launched for per user.
+The architecture for the PRAGMA cloud scheduler is shown in Figure~\ref{Fig:Arch}.
+
+In our pilot implementation of the cloud scheduler, we made a number of temporary assumptions to simplify the design.
+
+\begin{itemize}
+\item Only one reservation can be made per resource.
+While we made several changes in our pilot to customize the look of Booked, modifying it to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later version.
+\item Virtual cluster images are already available at each site.
+For our pilot, we worked with a small number of virtual cluster images and pre-installed them to each cloud system so they were readily accessible by pragma\_boot.
+We had the following three images: 1) a basic SGE Rocks virtual cluster, 2) an AutoDoc virtual cluster, and 3) a Lifemapper compute virtual cluster.
+\item Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.
+The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looked for reservations that are ready to be activated.
+Once it finds a reservation that need sto be activated, it launched pragma\_boot via SSH.
+It does not  monitor the virtual cluster nor email the user when their reservation is about to expire.
+\item Only single site virtual clusters can be launched.
+So far, we've only experimented with multi-site virtual clusters using IPOP and this requires the virtual cluster images to be pre-installed with IPOP and an automated configure script that will assign private IP addresses to each node on boot.
+For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.
+\end{itemize}
+
+cloud scheduler would leverage 
+To create a lightweight scheduler that enabled users to start and manage virtual clusters on leveraged the pragma\_boot, PCC, and software-defined networking tools that had been developed as part of PRAGMA
+As 
+Our 
+To schedule virtual clusters on the PRAGMA cloud testbed, 
 Discuss creation of Cloud Scheduler Pilot.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/arch.pdf}
+\caption{Architecture for the PRAGMA cloud scheduler.}
+\label{Fig:Arch}
+\end{center}
+\end{figure}
 

### 2015-08-14 paper.tex
Filled out more of pilot section and added screenshots

@@
 
-In our design discussions, we first looked at the existing solutions that could be adapted for PRAGMA's scheduling needs:
+In our design discussions, we first looked at some existing solutions that might be adapted for PRAGMA's scheduling needs.  We
 
@@
 
-The other key requirement for the cloud scheduler was that it should require minimal effort for a PRAGMA sites to participate in the PRAGMA testbed.
+The other key requirement for the cloud scheduler was that it should require minimal effort for PRAGMA sites to participate in the PRAGMA testbed.
 Therefore, we opted for a simple client-server architecture, where the client is a small PRAGMA package that interfaces to a site's local cloud system.
@@
 
+\textbf{Only one reservation can be made per resource:}  While we made several changes in our pilot to customize the look of Booked, modifying it to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later version.
+
+\textbf{Virtual cluster images are already available at each site:}  For our pilot, we worked with a small number of virtual cluster images and pre-installed them to each cloud system so they were readily accessible by pragma\_boot.
+We had the following three images: 1) a basic SGE Rocks virtual cluster, 2) an AutoDoc virtual cluster, and 3) a Lifemapper~\cite{lifemapper} compute virtual cluster.
+
+\textbf{Developed and used a PCC stub:}  Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.
+The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looked for reservations that are ready to be activated.
+Once it finds a reservation that need sto be activated, it launches pragma\_boot via SSH.
+It does not  monitor the virtual cluster nor email the user when their reservation is about to expire.
+
+\textbf{Only single site virtual clusters can be launched:}  So far, we've only experimented with multi-site virtual clusters using IPOP and this requires the virtual cluster images to be pre-installed with IPOP and an automated configure script that will assign private IP addresses to each node on boot.
+For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.
+
+Much of our work in this pilot focused on creating the PRAGMA Booking interface.  This required the following customizations to Booked:
+
 \begin{itemize}
-\item Only one reservation can be made per resource.
-While we made several changes in our pilot to customize the look of Booked, modifying it to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later version.
-\item Virtual cluster images are already available at each site.
-For our pilot, we worked with a small number of virtual cluster images and pre-installed them to each cloud system so they were readily accessible by pragma\_boot.
-We had the following three images: 1) a basic SGE Rocks virtual cluster, 2) an AutoDoc virtual cluster, and 3) a Lifemapper compute virtual cluster.
-\item Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.
-The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looked for reservations that are ready to be activated.
-Once it finds a reservation that need sto be activated, it launched pragma\_boot via SSH.
-It does not  monitor the virtual cluster nor email the user when their reservation is about to expire.
-\item Only single site virtual clusters can be launched.
-So far, we've only experimented with multi-site virtual clusters using IPOP and this requires the virtual cluster images to be pre-installed with IPOP and an automated configure script that will assign private IP addresses to each node on boot.
-For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.
+\item Added a custom field for a public SSH key text  to the user profile.
+When PCC launches a virtual cluster, it will pull the user's public SSH key from their profile and give it to pragma\_boot so that the user can login as root once the virtual cluster is launched.
+\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to reservations.
+\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT-enabled to resources.
+\item Added a numeric count as a custom field type.
+Booked only provided text matching on search so if a host had 16 CPUs available and the user requested 8 CPUs, the search would fail since "16" != "8".
+By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed since 8 <= 16.
+\item Added custom reservation statuses: "Starting", "Running", and "Stopping".  Each reservation status is shown as a different color in  the calendar view.
+\item Added the ability to retrieve and set the reservation status from the Booked REST API.  	
+\item Added the PRAGMA logo to the header.
 \end{itemize}
+
+We also fixed the following bugs in Booked and plan to work with the developer to integrate them back into the distribution:
+
+\begin{itemize}
+\item Fixed bug that allowed users from making reservations for past time frames.
+\item Fixed a time conversion bug that was preventing updates to existing reservations.
+\item Fixed bug where the username was getting set to a blank value when a reservation was updated via the REST API. 
+\item Fixed bug that would not recognize zero as a valid value (e.g., specifying a virtual cluster with just a frontend and 0 compute nodes).  
+\end{itemize}
+
+Some sample screenshots of the PRAGMA Booking user interface is shown in Figure~\ref{Fig:Booked}.
+Likewise, Figure~\ref{Fig:Reports} shows some screenshots of the administrations options and usage reports.
+Our customizations for Booked were packaged into a Rocks roll~\cite{cloudscheduler}, which automates its installation and publishes some sample data.
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/bookedReservationScreenshot.pdf}
+\caption{Screenshots of the PRAGMA Booking user interface showing how a user can view and filter resources, view reservation summaries, and create a new reservation.}
+\label{Fig:Booked}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/bookedReportScreenshot.pdf}
+\caption{Screenshots of the PRAGMA Booking admin and reporting interface showing where an administrator can edit configuration details and then view usage reports.}
+\label{Fig:Reports}
+\end{center}
+\end{figure*}
 

### 2015-08-14 paper.tex
Merging design

@@
 \begin{document}
-
+\sloppy %gets rid of word overrun in some columns
 \title{Experiences designing and building a PRAGMA Cloud Scheduler}
@@
 
-In our design discussions, we first looked at some existing solutions that might be adapted for PRAGMA's scheduling needs.  We
+In our design discussions, we first looked at some existing solutions that might be adapted for PRAGMA's scheduling needs.
+Open source batch schedulers like Slurm~\cite{slurm}, TORQUE~\cite{torque}, or HTCondor, are used to manage space shared resources.
+There are also other testbeds like Grid'5000~\cite{grid5000},  GENI~\cite{geni}, and PlanetLab~\cite{planetlab} that have their own tools for using and managing their infrastructure.
+In general, we found these tools had different goals and were too heavy weight for PRAGMA.
+Some tools were better documented than others but all had a fairly high learning curve and would have been non-trivial to adapt for PRAGMA's cloud testbed, which had a strong application and virtual cluster focus.
 
-\textbf{Existing HPC schedulers}:  Open source batch schedulers like Slurm~\cite{slurm}, TORQUE~\cite{torque}, and HTCondor are commonly used in  HPC to manage distributed  shared resources.
-While these batch schedulers  are widely used, they either do not work across multiple sites (TORQUE) or are too heavy weight for what PRAGMA needs, requiring custom configuration and multiple daemons at each site (Slurm and HTCondor).
-Furthermore, to be adopted by PRAGMA users who are scientists in many different domains,  we wanted a low barrier to entry like a simple Web based interface for scheduling, rather than the command-line tools and queue abstractions that these batch schedulers provide.
+We next looked more broadly for a simpler approach to scheduling and found a number of Web-based calendar reservation systems that seemed promising.
+These systems were designed to manage meeting room reservations, hotel room reservations, or equipment rentals.
+Their interfaces would be also be familiar to users in their everyday lives and thus well suited to the different PRAGMA users who have varying levels of distributed systems and networking expertise.
+Also, several of these tools were significantly less complex than the batch and testbed scheduling tools above.
+Later, we found a Web-based calendar reservation system called GridPrems~\cite{gridprems} for Grid'5000 indicating a simpler interface was sometimes needed even for their users.
 
-\textbf{Grid'5000:}   NEEDS TO BE WRITTEN.
-Add short description of Grid'5000.
-They use OAR which is a batch scheduler similar to above and the calendar gui called GridPrems is one of three options for interacting with Grid'5000.
-A key difference is that testbed resources are dedicated rather than like on PRAGMA where they could be shared;  I believe also single admin model.
-Interesting set of tools.
-We can say we liked their calendar approach the best.
+%Some of the calendar reservation systems we looked at like Reservations~\cite{drupalreservations}, Merci~\cite{merci}, and Booking\_calendar~\cite{wordpressbooking} were plugins for Drupal~\cite{drupal} or Wordpress~\cite{wordpresss}.
+However, the majority of tools were PHP based with a relational database backend like the Meeting Room Booking System~\cite{mrbs}, Booked~\cite{booked}, and Classroombookings~\cite{classroombookings}.
 
-\textbf{DHCP Leases:}  NEEDS TO BE WRITTEN.
-
-\textbf{GENI/PlanetLab}:  NEEDS TO BE WRITTEN.  ORCA
-
-Out of these approaches, we liked the Grid'5000 calendar reservation interface, GridPrems, for scheduling the best.
-The calendar reservation scheduling is an interface that all users are familiar with in their everyday lives and would be well suited to the different PRAGMA users who have varying levels of cloud expertise.
-
-\section{Calendar systems}
-\label{Sec:Calendars}
-
-Since we were primarily interested in a scheduler with a Web based calendar reservation interface like GridPrems, we decided to look at other open source calendar reservation systems that were available.
+Since we were primarily interested in a scheduler with a like GridPrems, we decided to look at other open source calendar reservation systems that were available.
 We looked at several systems used primarily for tracking room and/or equipment reservations and evaluated the following criteria:
@@
 However, we eliminated a number of tools because they had either too little documentation or complex installation instructions with several manual steps.
-
-We searched for candidate reservation tools using Google search and also searching specifically on Github and SourceForge.  The above criteria filtered out several tools.
-
-One category of tools we looked at were plugins to existing frameworks like Drupal~\cite{drupal} and Wordpress\~cite{wordpress}.
-We looked at Booking\_calendar for Wordpress as well as Merci and Reservations for Drupal
-One category of tools that we looked at were developed in PHP and had a relational backend database such as Postgres an MySQL.
-We looked at Meeting Room Booking System on SourceForge~\cite{mrbs}, Booked from Twinkle Toes Software, and Classroombookings on Github,
 
@@
 For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.
+Since IPOP is installed at the user-level, it can be used by any PRAGMA site but some sites may support better performing network overlays created with ViNE~\cite{vine} or software-defined networking tools.
+Therefore, our long-term goal is to automatically select the most appropriate network overlay depending on the sites that a user wants to include in their virtual cluster where IPOP is the lowest common denominator.
 

### 2015-08-14 paper.tex
Added last and first sections

@@
 PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
-Over the past four years, the technology focus for PRAGMA partners has shifted to cloud and software defined networking as enabling technologies.
-During PRAGMA 27 meeting, the Resources Working Group discussed rebooting the persistent PRAGMA Testbed as a way for sites to contribute and use shared resources, leveraging technologies such as PRAGMA Boot, Personal Cloud Controller (PCC), and virtual network overlays (i.e.
-IPOP and ViNe).
-To make PRAGMA resource sharing easier, a lightweight scheduler was proposed and discussed as a way to enable access to the resources and to manage resources reservations.
-This short paper discusses the design process and initial prototype of a simple cloud scheduler for PRAGMA.
+Over the past four years, the technology focus for PRAGMA's testbed has shifted to cloud and software defined networking as enabling technologies.
+This short paper describes the design of a simple cloud scheduler built on top of a Web reservation system that will enable users to easily run and manage virtual clusters for their science.
+It is also equally important that there be a low barrier to entry for PRAGMA institutions to become part of the PRAGMA cloud testbed and install the needed software.
+Therefore, the cloud scheduler  has a simple client-server architecture and will be built upon existing technologies such as pragma\_boot, Personal Cloud Controller,  virtual network overlays (e.g.,.
+IPOP), and  and also a Web reservation system.
+We discuss our experiences implementing our pilot cloud scheduler and then describe future work.
 
@@
 
-We next looked more broadly for a simpler approach to scheduling and found a number of Web-based calendar reservation systems that seemed promising.
+We next looked more broadly for a simpler approach to scheduling and found a number of open source Web-based calendar reservation systems that seemed promising.
 These systems were designed to manage meeting room reservations, hotel room reservations, or equipment rentals.
@@
 
-%Some of the calendar reservation systems we looked at like Reservations~\cite{drupalreservations}, Merci~\cite{merci}, and Booking\_calendar~\cite{wordpressbooking} were plugins for Drupal~\cite{drupal} or Wordpress~\cite{wordpresss}.
+Some of the calendar reservation systems we looked at like Reservations~\cite{drupalreservations}, Merci~\cite{merci}, and Booking\_calendar~\cite{wordpressbooking} were plugins for Drupal~\cite{drupal} or Wordpress~\cite{wordpress}.
 However, the majority of tools were PHP based with a relational database backend like the Meeting Room Booking System~\cite{mrbs}, Booked~\cite{booked}, and Classroombookings~\cite{classroombookings}.
+For each system, we evaluated 1) how easy it would be to manage resources, reservations, and users as well as add new fields and features, 2) the GUI interface with respect to intuitive menus/navigation as well as a clean, modern, and uncluttered look, and 3)  how easy it was to install and setup a demo instance.
+One deficiency that all systems had was that they only allowed one reservation per resource.
 
-Since we were primarily interested in a scheduler with a like GridPrems, we decided to look at other open source calendar reservation systems that were available.
-We looked at several systems used primarily for tracking room and/or equipment reservations and evaluated the following criteria:
+Out of the calendar reservation systems we evaluated, Booked~\cite{booked} from Twinkle Toes software seemed the best suited.
+It was easy to setup, had  a nice looking GUI interface, and was customizable.
+It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions, and user/group quotas.
+After working with Booked during our pilot, we did find some weaknesses.
+First, the documentation was a bit sparse in areas and any needed PHP changes were time consuming due to its heavily object-oriented structure (i.e., we had to go thru many files to find the right place to change the code).
 
-\textbf{Open Source and customizable:}  Open source software is generally preferable because it provides it is free to use and can be customized as needed.
-Since we did not know of any existing Web interfaces for managing virtual clusters, we limited our search to customizable Open Source tools and evaluated how easy it would be to manage resources and users as well as add new fields and features.
-
-\textbf{Nice GUI interface:}  We evaluated the GUI interfaces of several calendar reservation tools and looked for those that had intuitive menus and navigation and  a clean and uncluttered look.
-Some of the GUI interfaces we looked at just used basic HTML tags and looked very clunky so we limited our search to tools that leveraged CSS and Javascript and looked more professional.
-
-\textbf{Installation and maintenance:}  We downloaded a number of tools and reviewed their installation instructions to see how easy they were to setup a demo instance.
-Most of the tools we looked at were developed in PHP and had a relational database backend that was could be served from a basic  Apache web server\~cite{apache} setup.
-A few of the tools were plugins for Drupal.
-However, we eliminated a number of tools because they had either too little documentation or complex installation instructions with several manual steps.
-
-Discuss selection of Booked.
-
-Open source
-Easy to setup
-Nice GUI interface
-Report features
-REST API
-Customizable-ish
-LDAP and Active Directory support.
-Fine tuned roles and permissions.
-User and group quotas.
-
-Can only handle one reservation per resource at a time
-PHP changes can be painful (heavy OO makes it hard to find right files)
-Doc is sparse
-
-http://mrbs.sourceforge.net
-	PHP, Postgres/MySQL
-	Demo was a bit slow and a bit clunky looking
-Booked - http://www.twinkletoessoftware.com/products
-	Nice interface, lot of customization features
-	PHP, MySQL
-	RESTFUL API
-http://classroombookings.com/
-	Nice interface
-	PHP, MySQL
-http://www.unitime.org/unitime\_intro.php
-	Geeky looking, might have some good algorithms
-	Interface not as smooth
-	http://journal.code4lib.org/articles/2941
-	Clunky looking
-http://www.drupalrooms.com
-	More for hotel web sites ? e.g., prices, etc.
-https://groups.drupal.org/node/137544
-	Merci - https://www.drupal.org/project/merci
-		For equipment, etc.
-	https://www.drupal.org/project/reservations
-		Also seems pretty flexible
-
- tried a few tools  here is the summary:
-1. OTA hotel management   http://rocks-86.sdsc.edu/hotelmis/setup/initialsetup.php
-    appears to be just for the hotel, no easy login, no easy change. Not useful
-2. Mybooking - in Malay and after changing the language and trying enenglishigsh i did not get far. -- hotel
-3. Booking\_calendar - strictly room booking, very small in features, too much to rewrite. -- wordpress plugin
-4 Optaplanner appears to be a java app that runs outside of web. I tried examples and i am
-   not sure of its usability. The examples are running producing sometimes some output sometimes noe
-   and they are just java apps on some data. The examples  are not impressive even in terms of giving
-   an idea how things are run.
-
-A few more links are on http://opensource.com/business/15/1/top-project-management-tools-2015
-Started looking at  one of them, Tuleap, may be too big for what we need.
-
-\section{Cloud Scheduler Pilot}
-\label{Sec:Pilot}
+\section{Scheduler Architecture}
+\label{Sec:Arch}
 
@@
 The architecture for the PRAGMA cloud scheduler is shown in Figure~\ref{Fig:Arch}.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/arch.pdf}
+\caption{Architecture for the PRAGMA cloud scheduler.}
+\label{Fig:Arch}
+\end{center}
+\end{figure}
+
+\section{Scheduler Pilot}
+\label{Sec:Pilot}
 
@@
 
-cloud scheduler would leverage 
-To create a lightweight scheduler that enabled users to start and manage virtual clusters on leveraged the pragma\_boot, PCC, and software-defined networking tools that had been developed as part of PRAGMA
-As 
-Our 
-To schedule virtual clusters on the PRAGMA cloud testbed, 
-Discuss creation of Cloud Scheduler Pilot.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/arch.pdf}
-\caption{Architecture for the PRAGMA cloud scheduler.}
-\label{Fig:Arch}
-\end{center}
-\end{figure}
-
 \section{Conclusions and Future Work}
@@
 
-Some interesting conclusion.
+This paper described the the design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.
+Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation/maintenance overhead and opted for a simple client-server design.
+The server has an intuitive Web GUI frontend based on the Booked Web reservation system software.
+
+The pilot implementation of the cloud scheduler made a lot of temporary assumptions and so our future work addresses them.
+We will first enhance the Booked software to handle multiple reservations and then work on integrating IPOP and PRAGMA-ENT into the cloud scheduler as well.
+Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.
+Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging to each of the sites.
+Documentation and packaging will also be done throughout so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools..
 

### 2015-08-14 paper.tex
some text, update addresses, add some bbl

@@
 \title{Experiences designing and building a PRAGMA Cloud Scheduler}
-\subtitle{[Short Paper]}
 
@@
 Shava Smallen\\
-      \affaddr{San Diego Supercomputer Center}\\
-      \affaddr{University of California San Diego}\\
+      \affaddr{University of California, San Diego}\\
+      \affaddr{9500 Gilman Dr., MC 0505}\\
+      \affaddr{La Jolla, CA 92093, USA}\\
        \email{ssmallen@sdsc.edu}
@@
 Nadya Williams\\
-      \affaddr{San Diego Supercomputer Center}\\
-      \affaddr{University of California San Diego}\\
+      \affaddr{University of California, San Diego}\\
+      \affaddr{9500 Gilman Dr., MC 0505}\\
+      \affaddr{La Jolla, CA 92093, USA}\\
        \email{nadya@sdsc.edu}
@@
 \alignauthor Philip Papadopoulos \\
-      \affaddr{San Diego Supercomputer Center}\\
-      \affaddr{University of California San Diego}\\
+      \affaddr{University of California, San Diego}\\
+      \affaddr{9500 Gilman Dr., MC 0505}\\
+      \affaddr{La Jolla, CA 92093, USA}\\
        \email{phil@sdsc.edu}
@@
 \begin{abstract}
-PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
 Over the past four years, the technology focus for PRAGMA's testbed has shifted to cloud and software defined networking as enabling technologies.
-This short paper describes the design of a simple cloud scheduler built on top of a Web reservation system that will enable users to easily run and manage virtual clusters for their science.
-It is also equally important that there be a low barrier to entry for PRAGMA institutions to become part of the PRAGMA cloud testbed and install the needed software.
-Therefore, the cloud scheduler  has a simple client-server architecture and will be built upon existing technologies such as pragma\_boot, Personal Cloud Controller,  virtual network overlays (e.g.,.
-IPOP), and  and also a Web reservation system.
-We discuss our experiences implementing our pilot cloud scheduler and then describe future work.
+This short paper describes the design of a cloud scheduler reservation system  with a web interface that will enable users to easily run and manage virtual clusters for their science.
+It is equally important that there is a low barrier to entry for PRAGMA institutions to become part of the PRAGMA cloud testbed and install the needed software.
+Therefore, the cloud scheduler  has a simple client-server architecture and will be built upon existing technologies such as pragma\_boot, Personal Cloud Controller,  virtual network overlays (e.g., IPOP, ViNe), and a Web  interface for the reservation system.
+We discuss our implementation of the pilot cloud scheduler and then describe future work.
 
@@
 
-The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) researches actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
-Founded in 2002, PRAGMA originally sought to advance the use of grid technologies in applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.
+ PRAGMA researches actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology.
+Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rim.
 This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and to conduct scientific experiments.
-Testbed sites needed to install at a minimum Globus~\cite{globus} and a local HPC job scheduler as well as other optional software such as MPICH-G2~\cite{mpichg2} and Ninf-G~\cite{ninfg}.
+Testbed sites needed to install at a minimum Globus toolkit~\cite{globus}, a local HPC job scheduler as well as other optional software such as MPICH-G2 and Ninf-G.
 In November 2006, nineteen sites in thirteen countries were part of the testbed.
@@
 The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
-One early example was an AutoDock~\cite{autodock} virtual cluster created for the Avian Flu Grid research team in 2011~\cite{pragmaReport2011}.
-By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory, and 657 terabytes of online storage.
+One early example in 2011 was an AutoDock virtual cluster created for the Avian Flu Grid research team.
+By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory and 657 terabytes of online storage.
 In 2013 PRAGMA shifted its  focus to the creation and management of virtual clusters to make it easier for users to assemble a multi-node virtual environment for running their scientific experiments.
@@
 However, enabling a virtual cluster to be ported to different cloud hosting environment is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.
-In 2013, three pilot sites (UCSD, AIST, NCHC) developed a script for an automated virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
-Based on those experiences, the porting script was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.
-In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront~\cite{cloudfront}.
+In 2013, three pilot sites (UCSD, AIST, NCHC) developed a prototype for an automated virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
+Based on these experiences, the original prototype was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.
+In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront.
 
-\textbf{Personal Cloud Controller:}  to provide users with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014.
+\textbf{Personal Cloud Controller:}  to provide researchers with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014.
 This lightweight tool was designed to manage startup, status monitoring, and shutdown of a virtual cluster and was built on top of pragma\_boot and a well-known resource and workload management system HTCondor~\cite{condor}.
@@
 
-\textbf{Software-Defined Networking}:   PRAGMA began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network between multi-site virtual clusters and to protect access to sensitive datasets~\cite{pragmaReport2013}.
+\textbf{Software-Defined Networking}:   PRAGMA began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network among multi-site virtual clusters and to protect access to sensitive datasets.
 PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
@@
 
-In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate virtual hosts and clusters running on the different cloud deployments via leverage the above described technologies.
+In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate virtual hosts and clusters running on the different cloud deployments via leveraging the above described technologies.
 The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.
-A summary of calendar systems that were examined is provided in Section~\ref{Sec:Calendars}.
+A summary of calendar systems that were examined is provided in Section~\ref{Sec:Arch}.
 Section~\ref{Sec:Pilot} discusses our early implementation with one of the calendar systems called Booked and Section~\ref{Sec:Conclusions} concludes with our planned future work.
@@
 Open source batch schedulers like Slurm~\cite{slurm}, TORQUE~\cite{torque}, or HTCondor, are used to manage space shared resources.
-There are also other testbeds like Grid'5000~\cite{grid5000},  GENI~\cite{geni}, and PlanetLab~\cite{planetlab} that have their own tools for using and managing their infrastructure.
+There are also other testbeds like Grid'5000~\cite{grid5000},  GENI~\cite{geni}, and PlanetLab~\cite{planetlab} that have their own tools for using and managing their infrastructures.
 In general, we found these tools had different goals and were too heavy weight for PRAGMA.
@@
 
-We next looked more broadly for a simpler approach to scheduling and found a number of open source Web-based calendar reservation systems that seemed promising.
+We next searched for a simpler approach to scheduling and found a number of open source Web-based calendar reservation systems that seemed promising.
 These systems were designed to manage meeting room reservations, hotel room reservations, or equipment rentals.
-Their interfaces would be also be familiar to users in their everyday lives and thus well suited to the different PRAGMA users who have varying levels of distributed systems and networking expertise.
-Also, several of these tools were significantly less complex than the batch and testbed scheduling tools above.
+Their interfaces would be more familiar to users in their everyday lives and thus well suited to the different PRAGMA users who have varying levels of distributed systems and networking expertise.
+In addition, several of these tools were significantly less complex than the batch and testbed scheduling tools above.
 Later, we found a Web-based calendar reservation system called GridPrems~\cite{gridprems} for Grid'5000 indicating a simpler interface was sometimes needed even for their users.
@@
 However, the majority of tools were PHP based with a relational database backend like the Meeting Room Booking System~\cite{mrbs}, Booked~\cite{booked}, and Classroombookings~\cite{classroombookings}.
-For each system, we evaluated 1) how easy it would be to manage resources, reservations, and users as well as add new fields and features, 2) the GUI interface with respect to intuitive menus/navigation as well as a clean, modern, and uncluttered look, and 3)  how easy it was to install and setup a demo instance.
+For each system, we evaluated 1) how easy it would be to manage resources, reservations, and users as well as to add new parameters and features; 2) how intuitive the GUI interface was with respect to menus and navigation and if it had a clean, modern, and uncluttered look; 3)  how easfy it was to install and setup a prototype instance.
 One deficiency that all systems had was that they only allowed one reservation per resource.
 
-Out of the calendar reservation systems we evaluated, Booked~\cite{booked} from Twinkle Toes software seemed the best suited.
-It was easy to setup, had  a nice looking GUI interface, and was customizable.
-It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions, and user/group quotas.
-After working with Booked during our pilot, we did find some weaknesses.
-First, the documentation was a bit sparse in areas and any needed PHP changes were time consuming due to its heavily object-oriented structure (i.e., we had to go thru many files to find the right place to change the code).
+Booked reservation ~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compare to all the calendar reservation systems we evaluated.
+It was easy to setup, had  a good look and feel GUI interface, and was customizable.
+It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions management, and user and group quotas.
+After working with Booked during our pilot setup, we did find some weaknesses.
+First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (e.g., we had to go through many source code files to find the right place to change  or add functionality).
 
@@
 
-One key requirement for our cloud scheduler was to enable users to start and manage virtual clusters on the PRAGMA cloud testbed with minimal effort. 
+One key requirement for our cloud scheduler was to enable users to create and manage virtual clusters on the PRAGMA cloud testbed with minimal effort. 
 The basic workflow of how a user would start a virtual cluster and the steps are summarized below:
@@
 \item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
-\item The user selects the subset of resources they want, the timeframe, virtual cluster image, and configuration details and submits it to the Cloud scheduler.
+\item The user selects the subset of  desired resources, the timeframe, virtual cluster image, and configuration details and submits it to the Cloud scheduler.
 Once the reservation is verified and confirmed, the user will get an email confirmation.
-\item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when it's ready for them to login.
+\item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.
 \end{enumerate}
@@
 Figure~\ref{Fig:Flow} shows the above steps in a use case sequence diagram where \textit{PRAGMA Booking} refers to our customized Booked instance.
-PCC will monitor the health of the virtual cluster and when the reservation is close to expiring, will notify the user in case they want to extend their reservation.
-Otherwise, when the time expires or the user chooses to end their reservation, the virtual cluster will be shut down.
+PCC will monitor the health of the virtual cluster and when the reservation is close to expiring, will notify the user via email.
+The user then has an option to extend the reservation.
+Otherwise, when the time expires or the user chooses to end its reservation, the virtual cluster will be shut down.
 
@@
 The other key requirement for the cloud scheduler was that it should require minimal effort for PRAGMA sites to participate in the PRAGMA testbed.
-Therefore, we opted for a simple client-server architecture, where the client is a small PRAGMA package that interfaces to a site's local cloud system.
-Currently, the PRAGMA Package is just a SSH service and the pragma\_boot package, which is a small Python package that can be installed via RPM and can be configured in less than an hour.
-The server component is composed of the PRAGMA Booking interface and the PCC which gets launched for per user.
+Therefore, we opted for a simple client-server architecture, where the client is a small PRAGMA package that interfaces with a site's local cloud system.
+Currently, the PRAGMA Package consists of an SSH service and the pragma\_boot, which is a small Python package that can be installed via RPM and can be configured in less than an hour.
+The server component is composed of the PRAGMA Booking interface and the PCC which gets launched for each user.
 The architecture for the PRAGMA cloud scheduler is shown in Figure~\ref{Fig:Arch}.
@@
 
-\textbf{Only one reservation can be made per resource:}  While we made several changes in our pilot to customize the look of Booked, modifying it to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later version.
+\textbf{Only one reservation can be made per resource:}  While we made several changes in our pilot to customize the look of Booked, further modifications to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later time.
 
 \textbf{Virtual cluster images are already available at each site:}  For our pilot, we worked with a small number of virtual cluster images and pre-installed them to each cloud system so they were readily accessible by pragma\_boot.
-We had the following three images: 1) a basic SGE Rocks virtual cluster, 2) an AutoDoc virtual cluster, and 3) a Lifemapper~\cite{lifemapper} compute virtual cluster.
+We had the following three images: 1) a basic SGE Rocks virtual cluster; 2) an AutoDock virtual cluster;  3) a Lifemapper~\cite{lifemapper} compute virtual cluster.
 
 \textbf{Developed and used a PCC stub:}  Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.
-The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looked for reservations that are ready to be activated.
-Once it finds a reservation that need sto be activated, it launches pragma\_boot via SSH.
-It does not  monitor the virtual cluster nor email the user when their reservation is about to expire.
+The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looks for reservations that are ready to be activated.
+Once it finds a reservation that needs to be activated, it launches pragma\_boot via SSH.
+It does not  monitor the virtual cluster nor does it email the users when their reservation are about to expire.
 
-\textbf{Only single site virtual clusters can be launched:}  So far, we've only experimented with multi-site virtual clusters using IPOP and this requires the virtual cluster images to be pre-installed with IPOP and an automated configure script that will assign private IP addresses to each node on boot.
+\textbf{Only single site virtual clusters can be launched:}  So far, we have only experimented with multi-site virtual clusters using IPOP.
+This means that the IPOP  and an automated configure script that will assign private IP addresses to each node on boot are installed on the virtual cluster images.
 For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.
-Since IPOP is installed at the user-level, it can be used by any PRAGMA site but some sites may support better performing network overlays created with ViNE~\cite{vine} or software-defined networking tools.
-Therefore, our long-term goal is to automatically select the most appropriate network overlay depending on the sites that a user wants to include in their virtual cluster where IPOP is the lowest common denominator.
+IPOP can be used by any PRAGMA site but some sites may have support for better performing network overlays created with ViNE~\cite{vine} or software-defined networking tools.
+Therefore, our long-term goal is to automatically select the most appropriate network overlay depending on the  specification of the sites that a user wants to include in its virtual cluster and leave IPOP i usage as the lowest common denominator or a default.
 
@@
 \begin{itemize}
-\item Added a custom field for a public SSH key text  to the user profile.
-When PCC launches a virtual cluster, it will pull the user's public SSH key from their profile and give it to pragma\_boot so that the user can login as root once the virtual cluster is launched.
-\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to reservations.
-\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT-enabled to resources.
+\item Added a custom field for a public SSH key  to the user profile.
+When PCC launches a virtual cluster, it will pull the public SSH key from the user's  profile and give it to pragma\_boot so that the user can login as root once the virtual cluster is launched.
+The ability to use a root login is not required for all but can be needed in some case if users want to do system modifications.
+A user with a root login has a full ownership of the cluster.
+\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to reservations specifications.
+\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to resources. specifications.
 \item Added a numeric count as a custom field type.
-Booked only provided text matching on search so if a host had 16 CPUs available and the user requested 8 CPUs, the search would fail since "16" != "8".
-By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed since 8 <= 16.
+Booked only provided text matching on search so if a host had 16 CPUs available and the user requested 8 CPUs, the search would fail per "16" != "8" inequality.
+By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for 8 <= 16.
 \item Added custom reservation statuses: "Starting", "Running", and "Stopping".  Each reservation status is shown as a different color in  the calendar view.
@@
 \begin{itemize}
-\item Fixed bug that allowed users from making reservations for past time frames.
+\item Fixed bug that allowed users to make reservations for past time frames.
 \item Fixed a time conversion bug that was preventing updates to existing reservations.
@@
 
-Some sample screenshots of the PRAGMA Booking user interface is shown in Figure~\ref{Fig:Booked}.
+Some sample screenshots of the PRAGMA Booking user interface are shown in Figure~\ref{Fig:Booked}.
 Likewise, Figure~\ref{Fig:Reports} shows some screenshots of the administrations options and usage reports.
-Our customizations for Booked were packaged into a Rocks roll~\cite{cloudscheduler}, which automates its installation and publishes some sample data.
+We created a  Booked Rocks roll~\cite{cloudscheduler} which packages the software with our customizations and automates its installation and publishing of sample data.
 
@@
 This paper described the the design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.
-Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation/maintenance overhead and opted for a simple client-server design.
+Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation and maintenance overhead and opted for a simple client-server design.
 The server has an intuitive Web GUI frontend based on the Booked Web reservation system software.
 
-The pilot implementation of the cloud scheduler made a lot of temporary assumptions and so our future work addresses them.
-We will first enhance the Booked software to handle multiple reservations and then work on integrating IPOP and PRAGMA-ENT into the cloud scheduler as well.
+In the pilot implementation of the cloud scheduler  we made a lot of temporary assumptions, and  our future work will address them.
+We will first enhance the Booked software to handle multiple reservations and then work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.
 Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.
-Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging to each of the sites.
-Documentation and packaging will also be done throughout so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools..
+Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging them to each of the sites.
+We will package and document the resulting software so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools.
 

### 2015-08-14 paper.tex
Few small changes

--- 2015-08-14 paper.tex
@@
 \begin{abstract}
-PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
-Over the past four years, the technology focus for PRAGMA's testbed has shifted to cloud and software defined networking as enabling technologies.
-This short paper describes the design of a simple cloud scheduler built on top of a Web reservation system that will enable users to easily run and manage virtual clusters for their science.
+PRAGMA is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas such as  biodiversity distribution and lake ecology.
+PRAGMA's members collaborate on a testbed infrastructure and over the past four years, the technology focus has shifted to cloud and software defined networking as enabling technologies.
+This short paper describes the design of a simple cloud scheduler that will 1) enable users to easily run and manage virtual clusters for their science experiments, and 2) require minimal effort for PRAGMA members to participate in.
+.
 It is also equally important that there be a low barrier to entry for PRAGMA institutions to become part of the PRAGMA cloud testbed and install the needed software.
@@
 We discuss our experiences implementing our pilot cloud scheduler and then describe future work.
+built on top of a Web reservation system that will
 

### 2015-08-14 paper.tex

--- 2015-08-14 paper.tex
@@
 \sloppy %gets rid of word overrun in some columns
-\title{Experiences designing and building a PRAGMA Cloud Scheduler}
+\title{Lightweight Scheduling for the PRAGMA Cloud Testbed}
 
@@
 \begin{abstract}
-The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas of biodiversity distribution and lake ecology.
-Over the past four years, the technology focus for PRAGMA's testbed has shifted to cloud and software defined networking as enabling technologies.
-This short paper describes the design of a cloud scheduler reservation system  with a web interface that will enable users to easily run and manage virtual clusters for their science.
-It is equally important that there is a low barrier to entry for PRAGMA institutions to become part of the PRAGMA cloud testbed and install the needed software.
-Therefore, the cloud scheduler  has a simple client-server architecture and will be built upon existing technologies such as pragma\_boot, Personal Cloud Controller,  virtual network overlays (e.g., IPOP, ViNe), and a Web  interface for the reservation system.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) is a community of research scientists and institutions from around the Pacific Rim that work together to enable scientific expeditions in areas such as biodiversity distribution and lake ecology.
+PRAGMA's members collaborate on a testbed infrastructure and over the past four years, the technology focus has shifted to cloud and software defined networking as enabling technologies.
+This short paper describes the design of a web-based cloud scheduler reservation system that will enable users to easily run and manage virtual clusters for their science.
+Based on a client-server architecture, the cloud scheduler was designed so that it requires minimal installation and and management effort for participating PRAGMA testbed sites.
+The PRAGMA cloud scheduler is built using a web-based room reservation tool called Booked and leverages several PRAGMA technologies such as pragma\_boot, Personal Cloud Controller,  and virtual network overlays (e.g., IPOP, ViNe).
 We discuss our implementation of the pilot cloud scheduler and then describe future work.
-
 %add contribute and leverage shared resources
@@
 
- PRAGMA researches actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology.
-Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rim.
+The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) researchers actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
+Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rimm~\cite{pragmaReport2004}.
 This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and to conduct scientific experiments.
-Testbed sites needed to install at a minimum Globus toolkit~\cite{globus}, a local HPC job scheduler as well as other optional software such as MPICH-G2 and Ninf-G.
+Testbed sites needed to install at a minimum the Globus toolkit~\cite{globus}, a local HPC job scheduler as well as other optional software such as MPICH-G2~\cite{mpichg2} and Ninf-G~\cite{ninfg}.
 In November 2006, nineteen sites in thirteen countries were part of the testbed.
@@
 The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
-One early example in 2011 was an AutoDock virtual cluster created for the Avian Flu Grid research team.
+One early example in 2011 was an AutoDock virtual cluster created for the Avian Flu Grid research team~\cite{pragmaReport2011}.
 By 2012, the cloud testbed had ten sites with a total of 367 CPUs, 2.5 terabytes of memory and 657 terabytes of online storage.
-In 2013 PRAGMA shifted its  focus to the creation and management of virtual clusters to make it easier for users to assemble a multi-node virtual environment for running their scientific experiments.
+To make it easier for users to assemble a multi-node virtual environment for running their scientific experiments, PRAGMA shifted its  focus in 2013 to the creation and management of virtual clusters.
 The following technologies have been explored to facilitate the operation and use of virtual clusters on the PRAGMA cloud testbed.
 
-\textbf{Pragma\_boot:}  rather than requiring a single cloud system at all sites, PRAGMA encourages the sites to  deploy any virtualization technology that best fits their expertise.
-However, enabling a virtual cluster to be ported to different cloud hosting environment is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.
-In 2013, three pilot sites (UCSD, AIST, NCHC) developed a prototype for an automated virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
-Based on these experiences, the original prototype was redesigned and reimplemented as \textit{pragma\_boot} toolkit with "drivers" to support  virtual cluster migration to virtual environments enabled on Rocks and OpenNebula clusters.
-In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront.
+% Decide to capitalize first word since http://www.quickanddirtytips.com/education/grammar/formatting-vertical-lists?page=2 indicates it's easier to capitalize
+\textbf{Pragma\_boot:}  Rather than require a single cloud system at all sites, PRAGMA encourages sites to  deploy any virtualization technology that best fits their expertise.
+However, enabling a virtual cluster to be ported to different cloud hosting environments is a complex problem and requires tooling to retain the cluster relationship between head and worker nodes as well as software configurations.
+In 2013, three pilot sites (UCSD, AIST, NCHC) developed a script to automate virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
+Based on these experiences, the original prototype was redesigned and reimplemented as a \textit{pragma\_boot} toolkit with "drivers" to support  Rocks and OpenNebula.
+In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront~\cite{cloudfront}.
 
-\textbf{Personal Cloud Controller:}  to provide researchers with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014.
+\textbf{Personal Cloud Controller:}  To provide researchers with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014.
 This lightweight tool was designed to manage startup, status monitoring, and shutdown of a virtual cluster and was built on top of pragma\_boot and a well-known resource and workload management system HTCondor~\cite{condor}.
@@
 PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
-In 2014, The PRAGMA-ENT team worked to create an international reliable direct point-to-point data Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that allows to provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
+In 2014, The PRAGMA-ENT team worked to create an reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that allows to provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
 
-In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate virtual hosts and clusters running on the different cloud deployments via leveraging the above described technologies.
-The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered and why a simple calendar solution was selected.
-A summary of calendar systems that were examined is provided in Section~\ref{Sec:Arch}.
-Section~\ref{Sec:Pilot} discusses our early implementation with one of the calendar systems called Booked and Section~\ref{Sec:Conclusions} concludes with our planned future work.
+In late 2014 during the PRAGMA27 workshop, a lightweight scheduler was proposed to coordinate virtual hosts and clusters running on the different cloud deployments  and leveraging the above described technologies.
+The next section discusses the general requirements of the cloud scheduler and Section~\ref{Sec:Design} discusses the design options that were considered.
+The architecture of the cloud scheduler is provided in Section~\ref{Sec:Arch} and Section~\ref{Sec:Pilot} discusses our pilot implementation.
+Section~\ref{Sec:Conclusions} concludes with our planned future work.
 
@@
 
-PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users access to the resources at a  given time:
+PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users to access the resources at a  given time:
 
 \textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it had to be lightweight and required  minimal effort and minimal expertise for a site to add their cloud deployment to the list of resources available for scheduling.
-Similarly, the scheduler would need to work across multiple cloud deployments (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install specific cloud deployment tools in addition to their currently used.
+Similarly, the scheduler would need to work across multiple cloud systems (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install additional cloud deployment tools.
 
 \textbf{Easy to use}:  Currently, to execute a virtual cluster on the PRAGMA cloud testbed, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and launch their virtual cluster either using pragma\_boot  if it is available or manually via a sequence of defined steps.
-The design of the PRAGMA cloud scheduler should first simplify this process by ensuring that pragma\_boot gets deployed more broadly and port it to more cloud environments as needed.
-The scheduler should also provide a simple Web interface for users to see the available resources and sites, construct their virtual cluster, and manage their images.
+The design of the PRAGMA cloud scheduler should first simplify this process by ensuring that pragma\_boot gets deployed more broadly and port it to more cloud systems as needed.
+The scheduler should also provide a simple web interface for users to see the available resources and sites, construct their virtual cluster, and manage their images.
 
@@
 
-We next searched for a simpler approach to scheduling and found a number of open source Web-based calendar reservation systems that seemed promising.
+We next searched for a simpler approach to scheduling and found a number of open source web-based calendar reservation systems that seemed promising.
 These systems were designed to manage meeting room reservations, hotel room reservations, or equipment rentals.
-Their interfaces would be more familiar to users in their everyday lives and thus well suited to the different PRAGMA users who have varying levels of distributed systems and networking expertise.
+Their interfaces would be more familiar to users in their everyday lives and thus well suited to the different types of PRAGMA users.
 In addition, several of these tools were significantly less complex than the batch and testbed scheduling tools above.
-Later, we found a Web-based calendar reservation system called GridPrems~\cite{gridprems} for Grid'5000 indicating a simpler interface was sometimes needed even for their users.
+Later, we found a web-based calendar reservation system called GridPrems~\cite{gridprems} for Grid'5000 indicating a simpler interface was sometimes needed even for their users.
 
@@
 
-Booked reservation ~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compare to all the calendar reservation systems we evaluated.
-It was easy to setup, had  a good look and feel GUI interface, and was customizable.
+Booked~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compare to all the calendar reservation systems we evaluated.
+It was easy to setup, had  a good look and feel, and was customizable.
 It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions management, and user and group quotas.
@@
 One key requirement for our cloud scheduler was to enable users to create and manage virtual clusters on the PRAGMA cloud testbed with minimal effort. 
-The basic workflow of how a user would start a virtual cluster and the steps are summarized below:
+The basic workflow and steps for how a user would start a virtual cluster are summarized below:
 
 \begin{enumerate}
-\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (i.e., IPOP, ViNE, or PRAGMA-ENT) that is needed.  
+\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (i.e., IPOP, ViNE~\cite{vine}, or PRAGMA-ENT) that is needed.  
 \item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
@@
 Currently, the PRAGMA Package consists of an SSH service and the pragma\_boot, which is a small Python package that can be installed via RPM and can be configured in less than an hour.
-The server component is composed of the PRAGMA Booking interface and the PCC which gets launched for each user.
+The server component is composed of the PRAGMA Booking interface and PCC, which gets launched for each user.
 The architecture for the PRAGMA cloud scheduler is shown in Figure~\ref{Fig:Arch}.
@@
 \textbf{Developed and used a PCC stub:}  Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.
-The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and looks for reservations that are ready to be activated.
+The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and monitors reservations.
 Once it finds a reservation that needs to be activated, it launches pragma\_boot via SSH.
@@
 \textbf{Only single site virtual clusters can be launched:}  So far, we have only experimented with multi-site virtual clusters using IPOP.
-This means that the IPOP  and an automated configure script that will assign private IP addresses to each node on boot are installed on the virtual cluster images.
+When virtual cluster images have been IPOP-enabled, IPOP and an automated configure script will assign private IP addresses to each node on boot.
 For Rocks virtual clusters, there is a Rocks roll~\cite{ipoproll} that will automate the IPOP installation and install the configure scripts.
-IPOP can be used by any PRAGMA site but some sites may have support for better performing network overlays created with ViNE~\cite{vine} or software-defined networking tools.
-Therefore, our long-term goal is to automatically select the most appropriate network overlay depending on the  specification of the sites that a user wants to include in its virtual cluster and leave IPOP i usage as the lowest common denominator or a default.
+IPOP can be used by any PRAGMA site but some sites may have support for better performing network overlays created with ViNE or software-defined networking tools.
+Therefore, our long-term goal is to automatically select the most appropriate network overlay depending on the virtual cluster specification and leave IPOP as the lowest common denominator or default.
 
@@
 The ability to use a root login is not required for all but can be needed in some case if users want to do system modifications.
-A user with a root login has a full ownership of the cluster.
-\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to reservations specifications.
-\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to resources. specifications.
+A user with a root login has a full ownership of the virtual cluster.
+\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to the reservation specifications.
+\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to the resource specifications.
 \item Added a numeric count as a custom field type.
@@
 Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation and maintenance overhead and opted for a simple client-server design.
-The server has an intuitive Web GUI frontend based on the Booked Web reservation system software.
+The server has an intuitive web GUI frontend based on the Booked web reservation system software.
 

### 2015-08-14 paper.tex
Fixed small number of typos

@@
 The Pacific Rim Application and Grid Middleware Assembly (PRAGMA) researchers actively collaborate to enable scientific expeditions in areas of computational chemistry, telescience, biodiversity, and lake ecology~\cite{pragmaWeb}.
-Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rimm~\cite{pragmaReport2004}.
+Founded in 2002, PRAGMA originally sought to advance the use of grid technologies for applications among a community of investigators working with leading institution sites around the Pacific Rim~\cite{pragmaReport2004}.
 This included the deployment of a shared PRAGMA Grid testbed where participating sites could contribute and use resources as needed to develop and test new middleware and to conduct scientific experiments.
@@
 PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
-In 2014, The PRAGMA-ENT team worked to create an reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that allows to provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
+In 2014, The PRAGMA-ENT team worked to create an reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
 
@@
 However, the majority of tools were PHP based with a relational database backend like the Meeting Room Booking System~\cite{mrbs}, Booked~\cite{booked}, and Classroombookings~\cite{classroombookings}.
-For each system, we evaluated 1) how easy it would be to manage resources, reservations, and users as well as to add new parameters and features; 2) how intuitive the GUI interface was with respect to menus and navigation and if it had a clean, modern, and uncluttered look; 3)  how easfy it was to install and setup a prototype instance.
+For each system, we evaluated 1) how easy it would be to manage resources, reservations, and users as well as to add new parameters and features; 2) how intuitive the GUI interface was with respect to menus and navigation and if it had a clean, modern, and uncluttered look; 3)  how easy it was to install and setup a prototype instance.
 One deficiency that all systems had was that they only allowed one reservation per resource.
@@
 After working with Booked during our pilot setup, we did find some weaknesses.
-First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (e.g., we had to go through many source code files to find the right place to change  or add functionality).
+First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (i.e., we had to go through many source code files to find the right place to change  or add functionality).
 
@@
 Therefore, we opted for a simple client-server architecture, where the client is a small PRAGMA package that interfaces with a site's local cloud system.
-Currently, the PRAGMA Package consists of an SSH service and the pragma\_boot, which is a small Python package that can be installed via RPM and can be configured in less than an hour.
+Currently, the PRAGMA Package consists of an SSH service and pragma\_boot, which is a small Python package that can be installed via RPM and can be configured in less than an hour.
 The server component is composed of the PRAGMA Booking interface and PCC, which gets launched for each user.
@@
 A user with a root login has a full ownership of the virtual cluster.
-\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to the reservation specifications.
-\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to the resource specifications.
+\item Added custom fields for CPU count, amount of memory per host, and virtual cluster image name to the reservation specification.
+\item Added custom fields for CPU count, amount of memory per host, site hostname, and ENT capability (enabled, disabled) to the resource specification.
 \item Added a numeric count as a custom field type.

### 2015-08-15 paper.tex
update i.e reference

@@
 \begin{enumerate}
-\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (i.e., IPOP, ViNE~\cite{vine}, or PRAGMA-ENT) that is needed.  
+\item A user does a general search for available resources based on the number of CPUs, memory size, and networking configuration (e.g., IPOP, ViNE~\cite{vine}, or PRAGMA-ENT) that is needed.  
 \item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  

### 2015-09-24 paper.tex
Few mods to respond to reviewer comments

@@
 First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (i.e., we had to go through many source code files to find the right place to change  or add functionality).
+The granularity of Booked reservations is currently limited to an hourly basis and for today's PRAGMA needs this is sufficient because the time it takes to start up the virtual cluster
+is tens of minutes and  the user reservation is expected to  be on the order of days.
+The reservation granularity can be optimized in the future in order to avoid a potential of underutilized resources when the virtual cluster startup becomes minutes.
 
@@
 PCC will monitor the health of the virtual cluster and when the reservation is close to expiring, will notify the user via email.
-The user then has an option to extend the reservation.
-Otherwise, when the time expires or the user chooses to end its reservation, the virtual cluster will be shut down.
+The user then has an option to extend the reservation if resources are available.
+Otherwise, when the time expires or the user chooses to end the reservation, the virtual cluster will be shut down.
 

### 2016-01-13 paper.tex
Reformatted for journal submission and also made change to reflect that multiple reservations can be made per resource.

@@
-\documentclass{acm_proc_article-sp}
+\documentclass[times]{cpeauth}
 
+\usepackage{moreverb}
 \usepackage{url}
 
+\usepackage[dvips,colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
+
+\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
+T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\def\volumeyear{2010}
+
 \begin{document}
+
+\runningheads{A.~N.~Other}{A demonstration of the \journalabb\
+class file}
+
 \sloppy %gets rid of word overrun in some columns
+
 \title{Lightweight Scheduling for the PRAGMA Cloud Testbed}
 
-\numberofauthors{3}
+\author{Shava Smallen\affil{1}\corrauth, Nadya Williams\affil{1}, Philip Papadopoulos\affil{1}}
 
-\author{
-% 1st author      
-\alignauthor
-Shava Smallen\\
-      \affaddr{University of California, San Diego}\\
-      \affaddr{9500 Gilman Dr., MC 0505}\\
-      \affaddr{La Jolla, CA 92093, USA}\\
-       \email{ssmallen@sdsc.edu}
-% 2nd. author
-\alignauthor
-Nadya Williams\\
-      \affaddr{University of California, San Diego}\\
-      \affaddr{9500 Gilman Dr., MC 0505}\\
-      \affaddr{La Jolla, CA 92093, USA}\\
-       \email{nadya@sdsc.edu}
-\and % 3rd. author
-\alignauthor Philip Papadopoulos \\
-      \affaddr{University of California, San Diego}\\
-      \affaddr{9500 Gilman Dr., MC 0505}\\
-      \affaddr{La Jolla, CA 92093, USA}\\
-       \email{phil@sdsc.edu}
-%\and  % use '\and' if you need 'another row' of author names
-}
+\address{\affilnum{1}San Diego Supercomputer Center\break
+University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92093}\break
 
-\maketitle
+\corraddr{University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92093. E-mail: ssmallen@sdsc.edu}
+
 \begin{abstract}
@@
 
+\keywords{Cloud, Scheduling, Resource Sharing, Virtualization} % NOT required for Proceedings
+
+\maketitle
+
+\vspace{-6pt}
+
 % A category with the (minimum) three required fields
@@
 %terms{Cloud, Scheduling, Resource Sharing}
-
-\keywords{Cloud, Scheduling, Resource Sharing, Virtualization} % NOT required for Proceedings
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/flow.pdf}
+\includegraphics[width=\columnwidth]{figures/flow.ps}
 \caption{The flow of a user request to start a virtual cluster and the components that are involved in the cloud scheduler.}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/arch.pdf}
+\includegraphics[width=.75\columnwidth]{figures/arch.ps}
 \caption{Architecture for the PRAGMA cloud scheduler.}
@@
 In our pilot implementation of the cloud scheduler, we made a number of temporary assumptions to simplify the design.
-
-\textbf{Only one reservation can be made per resource:}  While we made several changes in our pilot to customize the look of Booked, further modifications to allow for multiple reservations per resource would have required deeper changes that we wanted to defer to a later time.
 
@@
 \begin{itemize}
+\item Added the ability to make multiple reservations on a resource per time slot to allow more than one virtual cluster to run simultaneously on a site.
+This was the most extensive change we made to Booked.
+Fortunately, no changes were needed to the backend database but it required several changes to the GUI code to allow an additional reservation to be made on a resource and to display multiple reservations per time slot in the calendar views.
+In total, enhancements were made to seven PHP files and two template files.
 \item Added a custom field for a public SSH key  to the user profile.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/bookedReservationScreenshot.pdf}
+\includegraphics[width=\textwidth]{figures/bookedReservationScreenshot.eps}
 \caption{Screenshots of the PRAGMA Booking user interface showing how a user can view and filter resources, view reservation summaries, and create a new reservation.}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/bookedReportScreenshot.pdf}
+\includegraphics[width=\textwidth]{figures/bookedReportScreenshot.eps}
 \caption{Screenshots of the PRAGMA Booking admin and reporting interface showing where an administrator can edit configuration details and then view usage reports.}
@@
 In the pilot implementation of the cloud scheduler  we made a lot of temporary assumptions, and  our future work will address them.
-We will first enhance the Booked software to handle multiple reservations and then work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.
+We will first work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.
 Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.

### 2016-07-07 paper.tex

@@
 
-\runningheads{A.~N.~Other}{A demonstration of the \journalabb\
-class file}
+\runningheads{A demonstration of the \journalabb\class file}
 
@@
 PRAGMA's members collaborate on a testbed infrastructure and over the past four years, the technology focus has shifted to cloud and software defined networking as enabling technologies.
-This short paper describes the design of a web-based cloud scheduler reservation system that will enable users to easily run and manage virtual clusters for their science.
+This paper describes the design of a web-based cloud scheduler reservation system that enables users to easily run and manage virtual clusters for their science.
 Based on a client-server architecture, the cloud scheduler was designed so that it requires minimal installation and and management effort for participating PRAGMA testbed sites.
@@
 PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
-In 2014, The PRAGMA-ENT team worked to create an reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
+In 2014, The PRAGMA-ENT team worked to create a reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
 
@@
 
-Booked~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compare to all the calendar reservation systems we evaluated.
+Booked~\cite{booked} from Twinkle Toes software seemed the best suited to our needs compared to all the calendar reservation systems we evaluated.
 It was easy to setup, had  a good look and feel, and was customizable.
@@
 After working with Booked during our pilot setup, we did find some weaknesses.
-First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (i.e., we had to go through many source code files to find the right place to change  or add functionality).
-The granularity of Booked reservations is currently limited to an hourly basis and for today's PRAGMA needs this is sufficient because the time it takes to start up the virtual cluster
-is tens of minutes and  the user reservation is expected to  be on the order of days.
-The reservation granularity can be optimized in the future in order to avoid a potential of underutilized resources when the virtual cluster startup becomes minutes.
+First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (for example, we had to go through many source code files to find the right place to change  or add functionality).
+Second, the reservation frequency of the event is currently limited to one hour.
+Based on today's PRAGMA needs this is sufficient because the time it takes to start up the virtual cluster
+is in the order of tens of minutes and  the user reservation is expected to be in the order of days.
+In the future when the virtual cluster startup  time shortens to minutes, the reservation frequency granularity can be optimized in order to avoid a potential of the underutilized resources.
 
@@
 \item Available resources that fit the criteria are displayed to the user and their availability is shown in a calendar layout.  
-\item The user selects the subset of  desired resources, the timeframe, virtual cluster image, and configuration details and submits it to the Cloud scheduler.
+\item The user selects the desired resources, the timeframe, virtual cluster image, and configuration details and submits them to the Cloud scheduler.
 Once the reservation is verified and confirmed, the user will get an email confirmation.
@@
 
-In our pilot implementation of the cloud scheduler, we made a number of temporary assumptions to simplify the design.
+In our pilot implementation of the cloud scheduler, we made a number of simplifying assumptions to the design more comprehensible.
 
@@
 
-\textbf{Developed and used a PCC stub:}  Since PCC is also under active development and was not easy to install, we used a simple stub to substitute for its functionality.
-The stub is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and monitors reservations.
+\textbf{Developed and used a PCC stub:}  To provide an interface for managing virtual cluster, we used a simple stub to emulate the functionality of PCC.
+The stub serves as a generic interface; it is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and monitors reservations.
 Once it finds a reservation that needs to be activated, it launches pragma\_boot via SSH.
@@
 \item Added a custom field for a public SSH key  to the user profile.
-When PCC launches a virtual cluster, it will pull the public SSH key from the user's  profile and give it to pragma\_boot so that the user can login as root once the virtual cluster is launched.
+When PCC launches a virtual cluster, it pulls the public SSH key from the user's  profile and gives it to pragma\_boot so that the user can login as root once the virtual cluster is launched.
 The ability to use a root login is not required for all but can be needed in some case if users want to do system modifications.
@@
 Booked only provided text matching on search so if a host had 16 CPUs available and the user requested 8 CPUs, the search would fail per "16" != "8" inequality.
-By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for 8 <= 16.
+By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for "8 \textless= 16".
 \item Added custom reservation statuses: "Starting", "Running", and "Stopping".  Each reservation status is shown as a different color in  the calendar view.

### 2016-07-14 paper.tex
Fix to get paper to build
Added a few more refs and few small grammar changes

@@
 
-\runningheads{A demonstration of the \journalabb\class file}
+%\runningheads{A demonstration of the \journalabb\class file}
 
@@
 In 2013, three pilot sites (UCSD, AIST, NCHC) developed a script to automate virtual cluster porting and demonstrated the same virtual cluster image being booted in three different cloud hosting environments, including Rocks/Xen, OpenNebula/KVM, and Amazon EC2.
-Based on these experiences, the original prototype was redesigned and reimplemented as a \textit{pragma\_boot} toolkit with "drivers" to support  Rocks and OpenNebula.
+Based on these experiences, the original prototype was redesigned and reimplemented as a \textit{pragma\_boot} toolkit with "drivers" to support  Rocks and OpenNebula~\cite{pragmaboot}.
 In 2014, a feature was added to pragma\_boot to download and boot virtual cluster images from Amazon CloudFront~\cite{cloudfront}.
 
-\textbf{Personal Cloud Controller:}  To provide researchers with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014.
+\textbf{Personal Cloud Controller:}  To provide researchers with an easy-to-use interface for managing the life cycle of their virtual clusters, a Personal Cloud Controller (PCC)  was started in 2014~\cite{pcc}.
 This lightweight tool was designed to manage startup, status monitoring, and shutdown of a virtual cluster and was built on top of pragma\_boot and a well-known resource and workload management system HTCondor~\cite{condor}.
@@
 After working with Booked during our pilot setup, we did find some weaknesses.
-First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure (for example, we had to go through many source code files to find the right place to change  or add functionality).
+First, the documentation was very sparse and any needed PHP changes were time consuming due to its heavily object-oriented structure.
+For example, we had to go through many source code files to find the right place to change  or add functionality.
 Second, the reservation frequency of the event is currently limited to one hour.
@@
 Once the reservation is verified and confirmed, the user will get an email confirmation.
+In the event two users try to reserve the same resources at the same time,
 \item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.
@@
 
-In our pilot implementation of the cloud scheduler, we made a number of simplifying assumptions to the design more comprehensible.
+In our pilot implementation of the cloud scheduler, we made a number of simplifying assumptions to the design.
 

### 2016-07-14 paper.tex
Add notes from call

@@
 
-\textbf{Developed and used a PCC stub:}  To provide an interface for managing virtual cluster, we used a simple stub to emulate the functionality of PCC.
+\textbf{Developed a stub for managing virtual clusters:}  To provide an interface for managing virtual clusters, we used a simple stub to emulate the functionality of PCC.
 The stub serves as a generic interface; it is a small Python script called \textit{pcc-check-reservations} that polls PRAGMA Booking via its REST interface and monitors reservations.
@@
 By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for "8 \textless= 16".
-\item Added custom reservation statuses: "Starting", "Running", and "Stopping".  Each reservation status is shown as a different color in  the calendar view.
+\item Added virtual cluster statuses: "Starting", "Running", and "Stopping".
+Each virtual cluster status is shown as a different color in  the calendar view.
+"Starting" means it has been approved and is being started up on the selected hypervisor...
+SHAVA EXPAND THIS
 \item Added the ability to retrieve and set the reservation status from the Booked REST API.  	
@@
 
-In the pilot implementation of the cloud scheduler  we made a lot of temporary assumptions, and  our future work will address them.
-We will first work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.
+In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.
+NADYA TO PUT SENTENCES FOR REVIEWER 1.
+In the future when we deploy on more sites, we can gather schedule statistics on usability and applicability of different scenarios.
+We will also work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.
 Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.

### 2016-07-15 paper.tex
More changes to address reviewer comments -- also added Nadya's change

@@
 It was easy to setup, had  a good look and feel, and was customizable.
-It also had some nice additional features like usage reporting, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions management, and user and group quotas.
+It also had some nice additional features like usage reporting, support for different time zones for each user, a REST API interface, LDAP and Active Directory support, fine grained roles and permissions management, and user and group quotas.
 After working with Booked during our pilot setup, we did find some weaknesses.
@@
 Once the reservation is verified and confirmed, the user will get an email confirmation.
-In the event two users try to reserve the same resources at the same time,
+Otherwise if there is a problem with the reservation such as someone else has taken the resources while the user was filling in the reservation fields, the user will get notified immediately of the failure once they hit the submit button or later via email.
 \item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.
@@
 Each virtual cluster status is shown as a different color in  the calendar view.
-"Starting" means it has been approved and is being started up on the selected hypervisor...
-SHAVA EXPAND THIS
+"Starting" means it has been accepted by the scheduler and is being started up on the local cloud systems, "Running" means that the resources are available for the user to log into, and "Stopping" means the virtual cluster is being shutdown and is no longer available for log in.
 \item Added the ability to retrieve and set the reservation status from the Booked REST API.  	
@@
 \end{itemize}
+
+We still need to address the scenario where two users might try to reserve the same resources at the same time as a race condition is possible.
+Booked does not have are not any locking mechanisms in place currently that would handle two simultaneous user requests for the same resources gracefully and so this will need to be added in the future.
+We also are not handling hardware failures at resource sites so when a  virtual cluster goes down; the failure would be recorded but we cannot accommodate relocation.
+For fault tolerance of the cloud scheduler, we will deploy a backup instance somewhere else so that we can failover in the event of an outage.
 
@@
 In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.
-NADYA TO PUT SENTENCES FOR REVIEWER 1.
-In the future when we deploy on more sites, we can gather schedule statistics on usability and applicability of different scenarios.
+Currently,  the startup latency from a virtual cluster reservation to its availability to the user has not been analyzed in detail but we observed it is on the order of tens of minutes for our pilot implementation.
+This user latency has multiple factors such as the underlying virtualization manager and load on the physical system where the virtual cluster is instantiated.
+There is additional latency if the cluster image is not available on the site as it will need to be downloaded from a remote repository and processed to make it available on the site.
+While this applies only for the first virtual cluster instantiation, it can add a considerable time  latency  depending on the cluster image size and the network speed from the site to the remote repository.
+
+In the future when we deploy on more sites, we can gather statistics on the cloud scheduler's usability and user experiences with virtual clusters startup latencies in addition to its applicability to different scenarios.
 We will also work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.

### 2016-07-16 paper.tex
reword some text

--- 2016-07-14 paper.tex
@@
 \item The user selects the desired resources, the timeframe, virtual cluster image, and configuration details and submits them to the Cloud scheduler.
-Once the reservation is verified and confirmed, the user will get an email confirmation.
-In the event two users try to reserve the same resources at the same time,
+Once the reservation is verified and confirmed, the user will get an email confirmation.  In the event two users try to reserve the same resource 
+the same time, for example, one user has taken the resources while the second user was filling in the reservation fields, the second user will get 
+a failure notification after hitting the submit button, and later via email.
 \item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.
@@
 By adding count as a custom field type, the search would do a numeric comparison on the values and the search would succeed for "8 \textless= 16".
-\item Added virtual cluster statuses: "Starting", "Running", and "Stopping".
-Each virtual cluster status is shown as a different color in  the calendar view.
-"Starting" means it has been approved and is being started up on the selected hypervisor...
-SHAVA EXPAND THIS
+\item Added virtual cluster statuses: "Starting", "Running", and "Stopping".  Each virtual cluster status is shown as a different color in  the
+calendar view.  "Starting" means the virtual cluster has been accepted by the scheduler and is being started up on the local cloud system, "Running"
+means that the virtual cluster is available for the user to log into, and "Stopping" means the virtual cluster is scheduled for shutdown and is no longer available for log in.
 \item Added the ability to retrieve and set the reservation status from the Booked REST API.  	
@@
 \end{itemize}
+
+We still need to address the race condition scenario where two users might try to reserve the same resources at the same time.  Booked currently 
+does not have any mechanisms in place that can handle two simultaneous user requests for the same resources gracefully. 
+The solution will need to be added in the future.  We also do not handle hardware failures at resource sites. When a virtual 
+cluster goes down the failure will be recorded but we cannot accommodate a cluster relocation.  For fault tolerance of the cloud scheduler, we will
+deploy a backup instance at a differnt site so that we can do a cloud scheduler server failover.
 
@@
 In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.  
-NADYA TO PUT SENTENCES FOR REVIEWER 1.
-In the future when we deploy on more sites, we can gather schedule statistics on usability and applicability of different scenarios.
-We will also work on integrating IPOP and PRAGMA-ENT into the cloud scheduler.
-Next, we will rework the PCC software and integrate HTCondor so it's used in personal mode allowing us to keep a light footprint on each of the PRAGMA sites.
-Finally, we will investigate using the CloudFront option in pragma\_boot to manage application virtual cluster images and staging them to each of the sites.
-We will package and document the resulting software so that new users and sites can get quickly onto the PRAGMA cloud testbed and use it as an experimental service to test new science and infrastructure tools.
+Currently, the virtual cluster startup latency from the reservation to its availability to the user has not been analyzed in detail; but we observed
+it is on the order of tens of minutes.   Multiple factors influence this latency, such as the underlying virtualization manager and load on the
+physical system where the virtual cluster is instantiated. There is additional latency if the cluster image is not available on the site as it will
+need to be downloaded from a remote repository and processed to make it available on the site.
+While this applies only to the first virtual cluster instantiation, it can add a considerable time  latency  depending on the cluster image size and the network speed from the site to the remote repository.
+
+In the future when we use deployment on more sites, we can gather statistics on the cloud scheduler's usability and user experiences with virtual
+clusters startup latencies in addition to its applicability to different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the
+cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it is used in personal mode allowing us to keep a light footprint on
+each of the PRAGMA sites.  Finally, we will investigate using the CloudFront option in pragma\_boot to manage application-specific  virtual cluster
+images and staging them to the sites.
+We will package and document the resulting software so that new users can quickly start using the PRAGMA cloud testbed as an experimental service to test new science and infrastructure tools.
 

### 2016-07-17 paper.tex
Few more tweaks to new sentences

@@
 Once the reservation is verified and confirmed, the user will get an email confirmation.  In the event two users try to reserve the same resource 
-the same time, for example, one user has taken the resources while the second user was filling in the reservation fields, the second user will get 
-a failure notification after hitting the submit button, and later via email.
+at the same time, for example, one user has taken the resources while the second user was filling in the reservation fields, the second user will get 
+a failure notification after hitting the submit button.
 \item When the reservation is ready to be activated, the cloud scheduler will use PCC and pragma\_boot to launch the virtual cluster, automatically configure the private and public network configuration on the virtual cluster nodes, and email the user when the reconfigured cluster is ready.
@@
 does not have any mechanisms in place that can handle two simultaneous user requests for the same resources gracefully. 
-The solution will need to be added in the future.  We also do not handle hardware failures at resource sites. When a virtual 
-cluster goes down the failure will be recorded but we cannot accommodate a cluster relocation.  For fault tolerance of the cloud scheduler, we will
-deploy a backup instance at a differnt site so that we can do a cloud scheduler server failover.
+We also do not handle hardware failures at resource sites. When a virtual 
+cluster goes down, the failure will be recorded but it will not be relocated to a different site.  We do plan to address the fault tolerance of the cloud scheduler by
+deploying a backup instance at a different site so that we can failover when needed.
 
@@
 In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.  
-Currently, the virtual cluster startup latency from the reservation to its availability to the user has not been analyzed in detail; but we observed
+Currently, the virtual cluster startup latency from the reservation to its availability to the user has not been analyzed in detail but we observed
 it is on the order of tens of minutes.   Multiple factors influence this latency, such as the underlying virtualization manager and load on the
@@
 
-In the future when we use deployment on more sites, we can gather statistics on the cloud scheduler's usability and user experiences with virtual
+In the future when the cloud scheduler is deployed to more sites, we can gather statistics on its usability and user experiences with virtual
 clusters startup latencies in addition to its applicability to different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the
 cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it is used in personal mode allowing us to keep a light footprint on
-each of the PRAGMA sites.  Finally, we will investigate using the CloudFront option in pragma\_boot to manage application-specific  virtual cluster
+each of the PRAGMA sites.  Finally, we will investigate using the remote repository option in pragma\_boot to manage application-specific  virtual cluster
 images and staging them to the sites.

### 2016-10-17 paper.tex
Add Beth and Renato's changes

@@
 This paper describes the design of a web-based cloud scheduler reservation system that enables users to easily run and manage virtual clusters for their science.
-Based on a client-server architecture, the cloud scheduler was designed so that it requires minimal installation and and management effort for participating PRAGMA testbed sites.
+Based on a client-server architecture, the cloud scheduler is designed to require minimal installation and management effort for participating PRAGMA testbed sites.
+%remove duplicate "and"; make present tense, simple substitution of enables
 The PRAGMA cloud scheduler is built using a web-based room reservation tool called Booked and leverages several PRAGMA technologies such as pragma\_boot, Personal Cloud Controller,  and virtual network overlays (e.g., IPOP, ViNe).
@@
 However, as of 2011, the number of sites started to decline and maintaining the numerous and complex middleware and scientific applications at each local site required significant people effort and expertise.
-Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and to lower the amount of effort any site would need to participate in the testbed.
+Since PRAGMA participants often have varying levels of of funding, staff, and expertise, PRAGMA started to shift away from grid towards cloud technologies to simplify the infrastructure and to lower the amount of effort any site would need to expend to
+%add "expend to"
+participate in the testbed.
 
-The first phase of the PRAGMA cloud testbed started with three sites and focused around the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
+The first phase of the PRAGMA cloud testbed started with three sites and focused on
+%replace "around" with "on"
+ the creation of application-specific virtual machines and making them portable to  different cloud hosting environments like Rocks~\cite{rocks}, OpenNebula~\cite{opennebula}, and Eucalyptus~\cite{eucalyptus}.
 One early example in 2011 was an AutoDock virtual cluster created for the Avian Flu Grid research team~\cite{pragmaReport2011}.
@@
 
-\textbf{Software-Defined Networking}:   PRAGMA began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network among multi-site virtual clusters and to protect access to sensitive datasets.
+\textbf{Software-Defined Networking}:   PRAGMA community members
+%add "community members"
+ began investigating software-defined networking technologies such as OpenFlow~\cite{openflow} in 2013 as a way to create a private network among multi-site virtual clusters and to protect access to sensitive datasets.
 PRAGMA then created the Experimental Network Testbed (PRAGMA-ENT) as a breakable, international testbed for use by PRAGMA researchers and collaborators.
-In 2014, The PRAGMA-ENT team worked to create a reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
+In 2014, the PRAGMA-ENT team worked to create a reliable international Layer-2 connection with network engineers and developed AutoVFlow~\cite{autovflow} that can provision private virtual network slices for each application, user, and/or project~\cite{pragmaReport2014}.
 
@@
 
-PRAGMA had the following three key requirements for a cloud testbed scheduler to enable multiple users to access the resources at a  given time:
+The PRAGMA resource providers
+%added "resource providers"
+ had the following three key requirements for a cloud testbed scheduler to enable multiple users to access the resources at a  given time:
 
-\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it had to be lightweight and required  minimal effort and minimal expertise for a site to add their cloud deployment to the list of resources available for scheduling.
+\textbf{Low participation overhead}:  One key requirement for the cloud testbed scheduler was that it 
+%remove "had to"; change to present tense
+be lightweight and require  minimal effort and minimal expertise for a site to add their cloud deployment to the list of resources available for scheduling.
 Similarly, the scheduler would need to work across multiple cloud systems (e.g., Rocks, OpenNebula, and OpenStack~\cite{openstack}) so that sites do not have to learn and install additional cloud deployment tools.
 
-\textbf{Easy to use}:  Currently, to execute a virtual cluster on the PRAGMA cloud testbed, users needs to contact each site individually to gain access to their cloud deployment, upload their images, and launch their virtual cluster either using pragma\_boot  if it is available or manually via a sequence of defined steps.
+\textbf{Easy to use}:  Currently, to execute a virtual cluster on the PRAGMA cloud testbed, users must
+%change "needs to" to "must"
+ contact each site individually to gain access to their cloud deployment, upload their images, and launch their virtual cluster either using pragma\_boot  if it is available or manually via a sequence of defined steps.
 The design of the PRAGMA cloud scheduler should first simplify this process by ensuring that pragma\_boot gets deployed more broadly and port it to more cloud systems as needed.
@@
 
-This paper described the the design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.
-Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use and low installation and maintenance overhead and opted for a simple client-server design.
-The server has an intuitive web GUI frontend based on the Booked web reservation system software.
+This paper describes the
+%present tense and remove extra "the"
+design and pilot implementation of a  scheduler for the PRAGMA cloud testbed.
+Since the scheduler does not need to scale beyond tens of users, we prioritized ease of use, low installation and maintenance overhead, and thus opted for a simple client-server design.
+%clarify what was prioritized
+The server has an intuitive web GUI front-end based on the Booked web reservation system software.
 
-In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address them.  
-Currently, the virtual cluster startup latency from the reservation to its availability to the user has not been analyzed in detail but we observed
-it is on the order of tens of minutes.   Multiple factors influence this latency, such as the underlying virtualization manager and load on the
+In the pilot implementation of the cloud scheduler  we made a lot of simplifying assumptions, and  our future work will address these.
+%these instead of them  
+Currently, the virtual cluster startup latency from the point in time of its reservation to its availability to the user has not been analyzed in detail but we observed
+it to be on the order of tens of minutes.   
+%clarify latency applied to
+Multiple factors influence this latency, such as the underlying virtualization manager and load on the
 physical system where the virtual cluster is instantiated. There is additional latency if the cluster image is not available on the site as it will
@@
 
-In the future when the cloud scheduler is deployed to more sites, we can gather statistics on its usability and user experiences with virtual
+Once
+%"once" instead of "in the future"
+ the cloud scheduler is deployed to more sites, we can gather statistics on its usability and user experiences with virtual
 clusters startup latencies in addition to its applicability to different scenarios.   We will also work on integrating IPOP and PRAGMA-ENT into the
-cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it is used in personal mode allowing us to keep a light footprint on
+cloud scheduler.  Next, we will rework the PCC software and integrate HTCondor so it is used in personal mode allowing us to keep a light footprint at
+%"at" instead of "on"
 each of the PRAGMA sites.  Finally, we will investigate using the remote repository option in pragma\_boot to manage application-specific  virtual cluster
